### 2024-04-11 paper/paper_cals.Rmd
Add five year age group comparisons for VMMC countries

---
title: "Circumcision Paper Text"
author: "Paddy O'Toole"
date: "2023-11-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- TODO: Change year from 2020 to 2023 -->

```{r, echo = FALSE}
# libraries
library(dplyr, warn.conflicts = FALSE)

# Metadata 
spec_agegroup <- "15-29"
spec_years <- c(2006, 2020)

ssa_iso3 <- sort(c(
  "AGO", "BDI", "BEN", "BFA", "BWA", "CAF", "CIV", "CMR", "COD",
  "COG", "ETH", "GAB", "GHA", "GIN", "GMB", "GNB", "GNQ", "KEN",
  "LBR", "LSO", "MLI", "MOZ", "MWI", "NAM", "NER", "NGA", "RWA",
  "SEN", "SLE", "SWZ", "TCD", "TGO", "TZA", "UGA", "ZAF", "ZMB", "ZWE"
))
```

```{r, echo = FALSE}
# Functions

# change numbers to be expressed in "millions"
to_mill <- \(x) round(x / 1e6, 3)

# change decimal to percentage
to_percentage <- \(x) paste0(round(x * 100, 2), "%")

# translate iso3 to country name
to_cntry_name <- function(x) {
  x_cntry = countrycode::countrycode(x, "iso3c", "country.name")
  # special cases
  case_when(
      x_cntry == "Congo - Brazzaville" ~ "Congo",
      x_cntry == "Congo - Kinshasa"    ~ "DR Congo",
      TRUE                             ~ x_cntry
  )
}

# simplify type
simplify_type <- function(dat) {
  dat %>% 
    mutate(type = case_when(
      grepl("MMC", type) ~ "MMC", 
      grepl("TMC", type) ~ "TMC", 
      TRUE               ~ "MC"
    ))
}
```

# Results

## Subnational Variation in Total, Medical \& Traditional Circumcision over time

```{r, echo = FALSE}
x1 <- readRDS("data/inlines/01_inlines.RDS")
for (i in seq_along(x1)) {
  assign(names(x1)[i], x1[[i]])
}
iso3_lowest_highest_cov <- simplify_type(iso3_lowest_highest_cov)
iso3_cov_diffs <- simplify_type(iso3_cov_diffs)
```

<!-- I surely need to add something about 10-29/15-29 year olds here? -->
<!-- Plot is of 15-29 year olds, but then I talk only about whole pop here -->
In 2006, the year before VMMC programme implementation began, an estimated 
`r n_circs[[1]] %>% filter(year == 2006, type == "MC") %>% pull(mean) %>% to_mill()` 
million men (95\% CI: 
`r n_circs[[1]] %>% filter(year == 2006, type == "MC") %>% pull(lower) %>% to_mill()` 
-
`r n_circs[[1]] %>% filter(year == 2006, type == "MC") %>% pull(upper) %>% to_mill()` 
million) were circumcised in the 33 countries in SSA for which we have valid 
circumcision data. 
Of these, 
`r n_circs[[1]] %>% filter(year == 2006, type == "MMC") %>% pull(mean) %>% to_mill()` 
million 
(`r n_circs[[1]] %>% filter(year == 2006, type == "MMC") %>% pull(lower) %>% to_mill()` 
- 
`r n_circs[[1]] %>% filter(year == 2006, type == "MMC") %>% pull(upper) %>% to_mill()` 
million) were MMCS, and 
`r n_circs[[1]] %>% filter(year == 2006, type == "TMC") %>% pull(mean) %>% to_mill()` 
million 
(`r n_circs[[1]] %>% filter(year == 2006, type == "TMC") %>% pull(lower) %>% to_mill()` 
- 
`r n_circs[[1]] %>% filter(year == 2006, type == "TMC") %>% pull(upper) %>% to_mill()` 
million) were TMCS. 
This contrasts with 2020, where an estimated 
`r n_circs[[1]] %>% filter(year == 2020, type == "MC") %>% pull(mean) %>% to_mill()` 
million males 
(`r n_circs[[1]] %>% filter(year == 2020, type == "MC") %>% pull(lower) %>% to_mill()` 
- 
`r n_circs[[1]] %>% filter(year == 2020, type == "MC") %>% pull(upper) %>% to_mill()` 
million) were circumcised in these same countries, of which 
`r n_circs[[1]] %>% filter(year == 2020, type == "MMC") %>% pull(mean) %>% to_mill()` 
million 
(`r n_circs[[1]] %>% filter(year == 2020, type == "MMC") %>% pull(lower) %>% to_mill()` 
- 
`r n_circs[[1]] %>% filter(year == 2020, type == "MMC") %>% pull(upper) %>% to_mill()` 
million) were performed as MMCS, and 
`r n_circs[[1]] %>% filter(year == 2020, type == "TMC") %>% pull(mean) %>% to_mill()` 
million 
(`r n_circs[[1]] %>% filter(year == 2020, type == "TMC") %>% pull(lower) %>% to_mill()` 
- 
`r n_circs[[1]] %>% filter(year == 2020, type == "TMC") %>% pull(upper) %>% to_mill()` 
million) as TMCS. 
This constituted an increase of 
`r n_circs[[2]] %>% filter(type == "MC") %>% pull(mean) %>% to_mill()` 
million 
(`r n_circs[[2]] %>% filter(type == "MC") %>% pull(lower) %>% to_mill()` 
- 
`r n_circs[[2]] %>% filter(type == "MC") %>% pull(upper) %>% to_mill()` 
million) newly circumcised males between 2006 and 2020, with 
`r n_circs[[2]] %>% filter(type == "MMC") %>% pull(mean) %>% to_mill()` 
million 
(`r n_circs[[2]] %>% filter(type == "MMC") %>% pull(lower) %>% to_mill()` 
- 
`r n_circs[[2]] %>% filter(type == "MMC") %>% pull(upper) %>% to_mill()` 
million) new MMCs performed, and 
`r n_circs[[2]] %>% filter(type == "TMC") %>% pull(mean) %>% to_mill()` 
million 
(`r n_circs[[2]] %>% filter(type == "TMC") %>% pull(lower) %>% to_mill()` 
- 
`r n_circs[[2]] %>% filter(type == "TMC") %>% pull(upper) %>% to_mill()` 
million) new TMCs performed. 
The disparity between the number of MCs performed and the number of MMCs performed can be explained by recalling that for five countries, Liberia, Senegal, Niger, Guinea and the Democratic Republic of the Congo, we have no circumcision type information, and so can only estimate MC, not MMC or TMC.
This translated to a change in mean MC coverage for the SSA region from 
`r covs[[1]] %>% filter(year == 2006, type == "MC") %>% pull(mean) %>% to_percentage()` 
(`r covs[[1]] %>% filter(year == 2006, type == "MC") %>% pull(upper) %>% to_percentage()` 
- 
`r covs[[1]] %>% filter(year == 2006, type == "MC") %>% pull(lower) %>% to_percentage()` ) 
in 2006 to 
`r covs[[1]] %>% filter(year == 2020, type == "MC") %>% pull(mean) %>% to_percentage()` 
(`r covs[[1]] %>% filter(year == 2020, type == "MC") %>% pull(upper) %>% to_percentage()` 
- 
`r covs[[1]] %>% filter(year == 2020, type == "MC") %>% pull(lower) %>% to_percentage()`)
in 2020, or an increase of
`r covs[[2]] %>% filter(type == "MC") %>% pull(mean) %>% to_percentage()` 
(`r covs[[2]] %>% filter(type == "MC") %>% pull(lower) %>% to_percentage()` 
- 
`r covs[[2]] %>% filter(type == "MC") %>% pull(upper) %>% to_percentage()`).
MMC coverage increased from 
`r covs[[1]] %>% filter(year == 2006, type == "MMC") %>% pull(mean) %>% to_percentage()` 
(`r covs[[1]] %>% filter(year == 2006, type == "MMC") %>% pull(lower) %>% to_percentage()` 
- 
`r covs[[1]] %>% filter(year == 2006, type == "MMC") %>% pull(upper) %>% to_percentage()` ) 
in 2006 to 
`r covs[[1]] %>% filter(year == 2020, type == "MMC") %>% pull(mean) %>% to_percentage()` 
(`r covs[[1]] %>% filter(year == 2020, type == "MMC") %>% pull(lower) %>% to_percentage()` 
- 
`r covs[[1]] %>% filter(year == 2020, type == "MMC") %>% pull(upper) %>% to_percentage()`) 
in 2020, an increase of 
`r covs[[2]] %>% filter(type == "MMC") %>% pull(mean) %>% to_percentage()` 
(`r covs[[2]] %>% filter(type == "MMC") %>% pull(lower) %>% to_percentage()` 
- 
`r covs[[2]] %>% filter(type == "MMC") %>% pull(upper) %>% to_percentage()`).
while TMC changed from 
`r covs[[1]] %>% filter(year == 2006, type == "TMC") %>% pull(mean) %>% to_percentage()` 
(`r covs[[1]] %>% filter(year == 2006, type == "TMC") %>% pull(lower) %>% to_percentage()` 
- 
`r covs[[1]] %>% filter(year == 2006, type == "TMC") %>% pull(upper) %>% to_percentage()`)
in 2006 to 
`r covs[[1]] %>% filter(year == 2020, type == "TMC") %>% pull(mean) %>% to_percentage()` 
(`r covs[[1]] %>% filter(year == 2020, type == "TMC") %>% pull(lower) %>% to_percentage()` 
- 
`r covs[[1]] %>% filter(year == 2020, type == "TMC") %>% pull(upper) %>% to_percentage()`)
in 2020, a decrease of
`r covs[[2]] %>% filter(type == "TMC") %>% mutate(mean = mean * -1) %>% pull(mean) %>% to_percentage()` 
(`r covs[[2]] %>% filter(type == "TMC") %>% mutate(upper = upper * -1) %>% pull(upper) %>% to_percentage()` 
- 
`r covs[[2]] %>% filter(type == "TMC") %>%  mutate(lower = lower * -1) %>%  pull(lower) %>% to_percentage()`).

<!-- TODO: Need to change this to be filtered by year as well!! -->
MC coverage in each country in 2006 ranged from 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2006) %>% slice(2) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2006) %>% slice(2) %>% pull(lower) %>% to_percentage()`
- 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2006) %>% slice(2) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2006) %>% slice(2) %>% pull(iso3) %>% to_cntry_name()`
to 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2006) %>% slice(1) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2006) %>% slice(1) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2006) %>% slice(1) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2006) %>% slice(1) %>% pull(iso3) %>% to_cntry_name()`, 
while MMC in 2006 ranged from 
`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2006) %>% slice(2) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2006) %>% slice(2) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2006) %>% slice(2) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2006) %>% slice(2) %>% pull(iso3) %>% to_cntry_name()`
to
`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2006) %>% slice(1) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2006) %>% slice(1) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2006) %>% slice(1) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2006) %>% slice(1) %>% pull(iso3) %>% to_cntry_name()`
and TMC ranged from 
`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2006) %>% slice(2) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2006) %>% slice(2) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2006) %>% slice(2) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2006) %>% slice(2) %>% pull(iso3) %>% to_cntry_name()`
to 
`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2006) %>% slice(1) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2006) %>% slice(1) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2006) %>% slice(1) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2006) %>% slice(1) %>% pull(iso3) %>% to_cntry_name()`.
In contrast, in 2020 MC coverage was estimated to have ranged from 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2020) %>% slice(2) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2020) %>% slice(2) %>% pull(lower) %>% to_percentage()`
- 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2020) %>% slice(2) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2020) %>% slice(2) %>% pull(iso3) %>% to_cntry_name()`
to 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2020) %>% slice(1) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2020) %>% slice(1) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2020) %>% slice(1) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2020) %>% slice(1) %>% pull(iso3) %>% to_cntry_name()`, 
while MMC ranged from 
`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2020) %>% slice(2) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2020) %>% slice(2) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2020) %>% slice(2) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2020) %>% slice(2) %>% pull(iso3) %>% to_cntry_name()`
to
`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2020) %>% slice(1) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2020) %>% slice(1) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2020) %>% slice(1) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2020) %>% slice(1) %>% pull(iso3) %>% to_cntry_name()`
and TMC ranged from 
`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2020) %>% slice(2) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2020) %>% slice(2) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2020) %>% slice(2) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2020) %>% slice(2) %>% pull(iso3) %>% to_cntry_name()`
to 
`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2020) %>% slice(1) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2020) %>% slice(1) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2020) %>% slice(1) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2020) %>% slice(1) %>% pull(iso3) %>% to_cntry_name()`.
<!-- Anything else to be said about this? -->

The largest percentage increase in MC coverage from 2006 to 2020 was 
in 
`r iso3_cov_diffs %>% filter(type == "MC") %>% slice(1) %>% pull(iso3) %>% to_cntry_name()`
at
`r iso3_cov_diffs %>% filter(type == "MC") %>% slice(1) %>% pull(mean) %>% to_percentage()`
(`r iso3_cov_diffs %>% filter(type == "MC") %>% slice(1) %>% pull(lower) %>% to_percentage()`
-
`r iso3_cov_diffs %>% filter(type == "MC") %>% slice(1) %>% pull(upper) %>% to_percentage()`),
from 
`r iso3_cov_2006_2020 %>% filter(type == "MC", year == 2006) %>% pull(mean) %>% to_percentage()`
(`r iso3_cov_2006_2020 %>% filter(type == "MC", year == 2006) %>% pull(lower) %>% to_percentage()`
-
`r iso3_cov_2006_2020 %>% filter(type == "MC", year == 2006) %>% pull(upper) %>% to_percentage()`)
in 2006 to 
`r iso3_cov_2006_2020 %>% filter(type == "MC", year == 2020) %>% pull(mean) %>% to_percentage()`
(`r iso3_cov_2006_2020 %>% filter(type == "MC", year == 2020) %>% pull(lower) %>% to_percentage()`
-
`r iso3_cov_2006_2020 %>% filter(type == "MC", year == 2020) %>% pull(upper) %>% to_percentage()`)
in 2020,
while the highest percentage increase in MMC was also in 
`r iso3_cov_diffs %>% filter(type == "MMC") %>% slice(1) %>% pull(iso3) %>% to_cntry_name()`
at
`r iso3_cov_diffs %>% filter(type == "MMC") %>% slice(1) %>% pull(mean) %>% to_percentage()`
(`r iso3_cov_diffs %>% filter(type == "MMC") %>% slice(1) %>% pull(lower) %>% to_percentage()`
-
`r iso3_cov_diffs %>% filter(type == "MMC") %>% slice(1) %>% pull(upper) %>% to_percentage()`),
from 
`r iso3_cov_2006_2020 %>% filter(type == "MMC", year == 2006) %>% pull(mean) %>% to_percentage()`
(`r iso3_cov_2006_2020 %>% filter(type == "MMC", year == 2006) %>% pull(lower) %>% to_percentage()`
-
`r iso3_cov_2006_2020 %>% filter(type == "MMC", year == 2006) %>% pull(upper) %>% to_percentage()`)
in 2006 to 
`r iso3_cov_2006_2020 %>% filter(type == "MMC", year == 2020) %>% pull(mean) %>% to_percentage()`
(`r iso3_cov_2006_2020 %>% filter(type == "MMC", year == 2020) %>% pull(lower) %>% to_percentage()`
-
`r iso3_cov_2006_2020 %>% filter(type == "MMC", year == 2020) %>% pull(upper) %>% to_percentage()`)
in 2020. 
For TMC, the largest percentage decrease between 2006 and 2020 was in 
`r iso3_cov_diffs %>% filter(type == "TMC") %>% slice(1) %>% pull(iso3) %>% to_cntry_name()`
at
`r iso3_cov_diffs %>% filter(type == "TMC") %>% slice(1) %>% pull(mean) %>% to_percentage()`
(`r iso3_cov_diffs %>% filter(type == "TMC") %>% slice(1) %>% pull(upper) %>% to_percentage()`
-
`r iso3_cov_diffs %>% filter(type == "TMC") %>% slice(1) %>% pull(lower) %>% to_percentage()`),
from 
`r iso3_cov_2006_2020 %>% filter(type == "TMC", year == 2006) %>% pull(mean) %>% to_percentage()`
(`r iso3_cov_2006_2020 %>% filter(type == "TMC", year == 2006) %>% pull(lower) %>% to_percentage()`
-
`r iso3_cov_2006_2020 %>% filter(type == "TMC", year == 2006) %>% pull(upper) %>% to_percentage()`)
in 2006 to 
`r iso3_cov_2006_2020 %>% filter(type == "TMC", year == 2020) %>% pull(mean) %>% to_percentage()`
(`r iso3_cov_2006_2020 %>% filter(type == "TMC", year == 2020) %>% pull(lower) %>% to_percentage()`
-
`r iso3_cov_2006_2020 %>% filter(type == "TMC", year == 2020) %>% pull(upper) %>% to_percentage()`)
in 2020.

The number of annual MCs performed increased from 
`r non_cum_n_performed %>% filter(type == "MC", year == 2006) %>% pull(mean) %>% to_mill()` 
million
(`r non_cum_n_performed %>% filter(type == "MC", year == 2006) %>% pull(lower) %>% to_mill()` 
-
`r non_cum_n_performed %>% filter(type == "MC", year == 2006) %>% pull(upper) %>% to_mill()` 
million) in 2006 to 
`r non_cum_n_performed %>% filter(type == "MC", year == 2006) %>% pull(mean) %>% to_mill()` 
million 
(`r non_cum_n_performed %>% filter(type == "MC", year == 2006) %>% pull(lower) %>% to_mill()` 
-
`r non_cum_n_performed %>% filter(type == "MC", year == 2006) %>% pull(upper) %>% to_mill()` 
million) in 2020, a growth of 
`r non_cum_n_performed_diff %>% filter(type == "MC") %>% pull(mean) %>% to_mill()` 
million (
`r non_cum_n_performed_diff %>% filter(type == "MC") %>% pull(lower) %>% to_mill()`
-
`r non_cum_n_performed_diff %>% filter(type == "MC") %>% pull(upper) %>% to_mill()` 
million) annually, averaged across all interceding years. 
Amongst these, 
`r non_cum_n_performed %>% filter(type == "MC", year == 2006) %>% pull(mean) %>% to_mill()` 
million
(`r non_cum_n_performed %>% filter(type == "MC", year == 2006) %>% pull(lower) %>% to_mill()` 
-
`r non_cum_n_performed %>% filter(type == "MC", year == 2006) %>% pull(upper) %>% to_mill()` 
million) were MMCs in 2006, while in 2020 
`r non_cum_n_performed %>% filter(type == "MC", year == 2020) %>% pull(mean) %>% to_mill()` 
million
(`r non_cum_n_performed %>% filter(type == "MC", year == 2020) %>% pull(lower) %>% to_mill()` 
-
`r non_cum_n_performed %>% filter(type == "MC", year == 2020) %>% pull(upper) %>% to_mill()` 
million) MMCs were performed, representing an increase in 
million 
(`r non_cum_n_performed_diff %>% filter(type == "MMC") %>% pull(lower) %>% to_mill()` 
-
`r non_cum_n_performed_diff %>% filter(type == "MMC") %>% pull(upper) %>% to_mill()` 
million) in MMCs performed annually. 
The number of TMCs performed annually did increase slightly from 2006, from 
`r non_cum_n_performed %>% filter(type == "TMC", year == 2006) %>% pull(mean) %>% to_mill()` 
million 
(
`r non_cum_n_performed %>% filter(type == "TMC", year == 2006) %>% pull(lower) %>% to_mill()` 
- 
`r non_cum_n_performed %>% filter(type == "TMC", year == 2006) %>% pull(upper) %>% to_mill()` 
million) to 
`r non_cum_n_performed %>% filter(type == "TMC", year == 2020) %>% pull(mean) %>% to_mill()`
million
(
`r non_cum_n_performed %>% filter(type == "TMC", year == 2020) %>% pull(lower) %>% to_mill()`
- 
`r non_cum_n_performed %>% filter(type == "TMC", year == 2020) %>% pull(upper) %>% to_mill()`
million) in 2020. However, 
TMC coverage actually decreased in many countries from 2006 to 2020. 
<!-- Should I give a number here? -->
This is the focus of section \ref{sec:org5fb5e18}.

<!-- *Second Plot* -->
<!-- Want to show here how ??? -->
Within countries, the median difference in MC coverage between districts with lowest and highest coverage in 2020 was `r district_diffs_vmmc$median_cov_diffs %>% filter(type == "MC") %>% pull(mean) %>% to_percentage()`
The smallest variation was in
`r district_diffs_vmmc$country_cov_diffs %>% filter(type == "MC") %>% pull(iso3) %>% last() %>% to_cntry_name()`, 
at 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MC") %>% slice(2) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MC") %>% slice(2) %>% pull(area_name)`
to
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MC") %>% slice(1) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MC") %>% slice(1) %>% pull(area_name)`
,
while the largest variation was in 
`r district_diffs_vmmc$country_cov_diffs %>% filter(type == "MC") %>% pull(iso3) %>% first() %>% to_cntry_name()`, 
at 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MC") %>% slice(4) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MC") %>% slice(4) %>% pull(area_name)`
to
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MC") %>% slice(3) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MC") %>% slice(3) %>% pull(area_name)`.
The median difference within countries in MMC coverage between districts with lowest and highest coverage in 2020 was `r district_diffs_vmmc$median_cov_diffs %>% filter(type == "MMC") %>% pull(mean) %>% to_percentage()`
The smallest variation was in
`r district_diffs_vmmc$country_cov_diffs %>% filter(type == "MMC") %>% pull(iso3) %>% last() %>% to_cntry_name()`, 
at
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MMC") %>% slice(2) %>% pull(mean) %>% to_percentage()`
\%
in
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MMC") %>% slice(2) %>% pull(area_name)`
to
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MMC") %>% slice(1) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MMC") %>% slice(1) %>% pull(area_name)`,
while the largest variation was in 
`r district_diffs_vmmc$country_cov_diffs %>% filter(type == "MMC") %>% pull(iso3) %>% first() %>% to_cntry_name()`, 
at 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MMC") %>% slice(4) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MMC") %>% slice(4) %>% pull(area_name)`
to
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MMC") %>% slice(3) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MMC") %>% slice(3) %>% pull(area_name)`.
For TMC coverage, the median difference between districts in 2020 was 
`r district_diffs_vmmc$median_cov_diffs %>% filter(type == "TMC") %>% pull(mean) %>% to_percentage()`
The smallest variation was in
`r district_diffs_vmmc$country_cov_diffs %>% filter(type == "TMC") %>% pull(iso3) %>% last() %>% to_cntry_name()`, 
at
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "TMC") %>% slice(2) %>% pull(mean) %>% to_percentage()`
\%
in
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "TMC") %>% slice(2) %>% pull(area_name)`
to
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "TMC") %>% slice(1) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "TMC") %>% slice(1) %>% pull(area_name)`,
while the largest variation was in 
`r district_diffs_vmmc$country_cov_diffs %>% filter(type == "TMC") %>% pull(iso3) %>% first() %>% to_cntry_name()`, 
at 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "TMC") %>% slice(4) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "TMC") %>% slice(4) %>% pull(area_name)`
to
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "TMC") %>% slice(3) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "TMC") %>% slice(3) %>% pull(area_name)`.
<!-- Anything else? --->

## Subnational Variation in total, medical \& traditional circumcision across different ages

```{r, echo = FALSE}
x2 <- readRDS("data/inlines/02_inlines.RDS")
for (i in seq_along(x2)) {
  assign(names(x2)[i], x2[[i]])
}
age_group_cov <- sapply(
  age_group_cov, simplify_type, simplify = FALSE, USE.NAMES = TRUE
)
age_group_cov_non_vmmc <- sapply(
  age_group_cov_non_vmmc, simplify_type, simplify = FALSE, USE.NAMES = TRUE
)
age_group_cov_vmmc <- sapply(
  age_group_cov_vmmc, simplify_type, simplify = FALSE, USE.NAMES = TRUE
)
```

<!-- TODO: Replace WCA/ECA with VMMC/Non-VMMC -->

Age at circumcision differed greatly by circumcision type and location. 
The five-year age group in 2006, the year before VMMC programme implementation began, 
which had the highest MC coverage across the SSA region was
`r age_group_cov$fyag_highest_covs %>% filter(type == "MC", year == 2006) %>% pull(age_group)`, 
at 
`r age_group_cov$fyag_highest_covs %>% filter(type == "MC", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$fyag_highest_covs %>% filter(type == "MC", year == 2006) %>% pull(lower) %>% to_percentage()`
-
`r age_group_cov$fyag_highest_covs %>% filter(type == "MC", year == 2006) %>% pull(upper) %>% to_percentage()`
)
. 
In contrast, the age group with the highest MC coverage in 2020 was 
`r age_group_cov$fyag_highest_covs %>% filter(type == "MC", year == 2020) %>% pull(age_group)`,
at
`r age_group_cov$fyag_highest_covs %>% filter(type == "MC", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$fyag_highest_covs %>% filter(type == "MC", year == 2020) %>% pull(lower) %>% to_percentage()`
-
`r age_group_cov$fyag_highest_covs %>% filter(type == "MC", year == 2020) %>% pull(upper) %>% to_percentage()`
).
The five-year age group in 2006 which had the highest MMC coverage was
`r age_group_cov$fyag_highest_covs %>% filter(type == "MMC", year == 2006) %>% pull(age_group)`, 
at 
`r age_group_cov$fyag_highest_covs %>% filter(type == "MMC", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$fyag_highest_covs %>% filter(type == "MMC", year == 2006) %>% pull(lower) %>% to_percentage()`
-
`r age_group_cov$fyag_highest_covs %>% filter(type == "MMC", year == 2006) %>% pull(upper) %>% to_percentage()`
)
. 
The age group with the highest MMC coverage in 2020 was also
`r age_group_cov$fyag_highest_covs %>% filter(type == "MMC", year == 2020) %>% pull(age_group)`,
, now at
`r age_group_cov$fyag_highest_covs %>% filter(type == "MMC", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$fyag_highest_covs %>% filter(type == "MMC", year == 2020) %>% pull(lower) %>% to_percentage()`
-
`r age_group_cov$fyag_highest_covs %>% filter(type == "MMC", year == 2020) %>% pull(upper) %>% to_percentage()`
).
The five-year age group in 2006 which had the highest TMC coverage was
`r age_group_cov$fyag_highest_covs %>% filter(type == "TMC", year == 2006) %>% pull(age_group)`, 
at 
`r age_group_cov$fyag_highest_covs %>% filter(type == "TMC", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$fyag_highest_covs %>% filter(type == "TMC", year == 2006) %>% pull(lower) %>% to_percentage()`
-
`r age_group_cov$fyag_highest_covs %>% filter(type == "TMC", year == 2006) %>% pull(upper) %>% to_percentage()`
)
. 
The age group with the highest TMC coverage in 2020 was
`r age_group_cov$fyag_highest_covs %>% filter(type == "TMC", year == 2020) %>% pull(age_group)`,
,
`r age_group_cov$fyag_highest_covs %>% filter(type == "TMC", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$fyag_highest_covs %>% filter(type == "TMC", year == 2020) %>% pull(lower) %>% to_percentage()`
-
`r age_group_cov$fyag_highest_covs %>% filter(type == "TMC", year == 2020) %>% pull(upper) %>% to_percentage()`
).
This decrease in the five-year age group with the highest MC coverage, increase in the 
age group for TMC with the highest coverage (as well as a decrease in coverage for 
the highest five-year age group), and increase in overall MMC coverage across the SSA region is 
indicative of the effect of VMMC programme implementation on the region.

<!-- TODO Still: 
- 15-29 year olds increase in MC/TMC/MMC (already written about?)
- Contrast to 30-49 year olds somehow, showing different patterns of circ
(increasing more for MC/MMC, decreasing more for TMC)
- Compare differences in VMMC and non-VMMC, in terms of the above
-->
We may compare changes in circumcision amongst two age groups; 15-29 year olds, which 
are a particular focus for VMMC implementation, and 30-49 year olds, which are not. 
MC coverage amongst 15-29 year olds increased from 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020. 
In contrast, MC coverage amongst 30-49 year olds only increased from 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020. 
Therefore, MC coverage across SSA saw greater increases in this younger age group than in this older one, in line 
with VMMC programme age focus. 
Similarly, MMC coverage amongst 15-29 year olds increased from 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020, while MMC coverage amongst 30-49 year olds increased from 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020. 
Therefore, MMC coverage across SSA also saw greater increases amongst 15-29 year olds 
than amongst 30-49 year olds. 
The changes seen in TMC, however, contrast to these larger increases in MC and MMC coverage for 15-29 year olds. 
TMC coverage amongst 15-29 year olds decreased from 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020, while TMC coverage amongst 30-49 year olds decreased from 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020. 
Thus, TMC coverage decreased less in absolute terms amongst 30-49 year olds than amongst 
15-29 year olds, which is line with the greater focus on performing MMCs amongst this 
younger age group.

Finally, we can contrast patterns in circumcision pattern for different age groups
amongst VMMC and non-VMMC priority countries, where differences in 2006 and 2020 
levels and patterns of circumcision seem to be more pronounced.

<!-- For VMMC countries, five year age group MC, MMC, TMC -->
For VMMC priority countries, five-year age group MC coverage in both 2006 and 2020 was highest amongst 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MC") %>% pull(age_group)`
year olds, at 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MC") %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MC") %>% pull(lower) %>% to_percentage()`
 - 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MC") %>% pull(upper) %>% to_percentage()`
) and 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MC") %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MC") %>% pull(lower) %>% to_percentage()`
 - 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MC") %>% pull(upper) %>% to_percentage()`
), respectively. 
However, in 2006 MMC coverage was highest amongst 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MMC") %>% pull(age_group)`
 year olds, at 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MMC") %>% pull(mean) %>% to_percentage()`
 (
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MMC") %>% pull(lower) %>% to_percentage()`
 - 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MMC") %>% pull(upper) %>% to_percentage()`
). 
This contrasts with 2020, where MMC coverage was highest amongst 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MMC") %>% pull(age_group)`
 year olds, at
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MMC") %>% pull(mean) %>% to_percentage()`
 (
x
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MMC") %>% pull(lower) %>% to_percentage()`
 - 
x
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MMC") %>% pull(upper) %>% to_percentage()`
), showing a shift upwards in age for MMCs in VMMC priority countries, as a result of VMMC focus on young adults. 
In 2006, TMC coverage was greatest in 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(age_group)`
year olds, at 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(mean) %>% to_percentage()`
 (
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(lower) %>% to_percentage()`
 - 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(upper) %>% to_percentage()`
). while in 2020 it was most prevalent in
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(age_group)`
year olds, at 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(mean) %>% to_percentage()`
 (
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(lower) %>% to_percentage()`
 - 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(upper) %>% to_percentage()`, highlighting a shift upwards in TMCs in VMMC priority countries, 
likely resulting from previously TMCs being performed as TMCs.

For non-VMMC countries, five year age group MC, MMC, TMC

Similarly, we can compare the 2006 and 2020 circumcision levels amongst 15-29 and 30-49 year olds
in both VMMC and non-VMMC countries, respectively.

## Subnational Heterogeneity in Age at Medical and Traditional Circumcision

Fill in! 
- Talk mainly about average age at which MCs/MMCs/TMCs performed in 2006 and 2020
- Also compare VMMC and non-VMMC

The mean age at circumcision for MC in 2006 was 
x
(
x
- 
x
), while in 2020 it was
x 
(
x
-
x
).
Similarly, the mean age at circumcision for MMC in 2006 was 
x
(
x
- 
x
), while in 2020 it was
x 
(
x
-
x
).
Finally, the mean age at circumcision in 2006 for TMC was 
x
(
x
- 
x
), while in 2020 it was
x 
(
x
-
x
).
So the mean age at circumcision for MC and MMC went up slightly from 2006 to 2020. 
However, it is important to note that the uncertainty ranges in our credible intervals 
overlap, so this increase is ?. 
In constant, the mean age at circumcision for TMC decreased. 
This may suggest that VMMC implementation amongst 15-29 year olds has 
corresponded to an upward shift of mean age at circumcision for MC and MMC, and 
replaced some circumcisions which would previously have been performed as TMCs, 
which may explain the observed decrease in mean age of TMC.

What is the narrative here?
- For SSA, MC and MMC age goes up, TMC goes down
- For VMMC, average age for all substantially higher, but follows same pattern. 
  Could be argued that this is because people previously receiving TMCs are now receiving MMCs, 
  pushing mean age of TMC down further. 
- For non-VMMC, ages substantially lower. 
  MMC age actually decreases, suggesting MMC-nT occuring, replacing TMC!

## Decrease in TMC in SSA region Over Time

The average TMC coverage in SSA in 2006 was 
x
\% (
x
\% to 
x
\%). 
This fell to 
x
\% (
x
\% to 
x
\%) in 2020, representing a decrease of 
x
\% (
x
\% to 
x
\%). 
In ESA, TMC coverage was 
x
\% (
x
\% to 
x
\%) in 2006, compared to 
x
\% (
x
\% to 
x
\%) in 2020, a 
x
\% (
x
\% to 
x
\%) decrease. 
In constrast, TMC coverage in WCA was 
x
\% (
x
\% to 
x
\%) versus 
x
\% (
x
\% to 
x
\%) in 2020, a 
x
\% (
x
\% to 
x
\%) fall. 
In 2006, 
x 
(
x 
to 
x
) TMCs were performed, while in 2020 this was reduced to 
x 
(
x to 
x
), representing a 
x (
x
to
x
) decrease in annual TMCs.
In ESA, the annual number of TMCs performed dropped from 
x
(
x
to 
x
) in 2006 to 
x
(
x
to 
x
) in 2020, a 
x 
(
x
to 
x
) drop. 
In WSA, this number fell from 
x 
(
x 
to 
x
) in 2006 to 
x 
(
x
to 
x
) in 2020, a fall of 
x
(
x
to 
x
).
<!-- TODO: Change to 15-29, right? -->
For the 10-29 age group, TMC coverage was reduced from 
x
\% (
x
\% to 
x
\%) in 2006 to 
x
\% (
x
\% to 
x
\%) in 2020, a drop of 
x
\% 
(
x
\% to 
x
\%).
In VMMC countries, this was reduced from 
x
\% (
x
\% to 
x
\%) to 
x
\% (
x
\% to 
x
\%) from 2006 to 2020, while even in non-VMMC countries this decreased from 
x
\% (
x
\% to 
x
\%) to 
x
\% (
x
\% to 
x
\%). 
The largest decrease in TMC coverage from 2006 to 2020 was in 
x
, from 
x
\% (
x
\% to 
x
\%) to 
x
\% (
 x 
\% to 
 x 
\%). 
[If the above is a VMMC country] The largest decrease in TMC coverage amongst non-VMMC countries was from 
x 
\% (
x 
\% to 
x 
\%) to 
x 
\% (
x 
\% to 
x 
\%) in 
x 
.

## UNAIDS Goals Progress amongst 10-29 Year Olds in VMMC Target Countries

Amongst the 13 VMMC priority countries modelled, just 
x
 countries were projected to have achieved the VMMC target of 90\% MC coverage amongst 10-29 year olds, 
x
 and 
x
. 
MC coverage across these countries increased from 
x
\% (
x
\% to 
x
\%) in 2006 to 
x
\% (
x
\% to 
x
\%) in 2020.

x
 (
x
 to 
x
) circumcisions were performed between 2007 and 2020 in these countries. 
Of these, 
x
 (
x
 to 
x
) were MMCs. 
No countries were e
x
pected to have reached this goal in all districts. 
The number of annual MCs increased from 
x
 (
x
 to 
x
) in 2006 to 
x
 (
x
 to 
x
) in 2020.

x
 (
x
 to 
x
) additional circumcisions are required to achieve 90\% MC coverage in all VMMC countries, 
x
 (
x
 to 
x
) times the projected number of MCs for 2020. 
The largest increase in MC coverage amongst VMMC priority countries was in 
x
, from 
x
\% (
x
\% to 
x
\%) in 2006 to 
x
\% (
x
\% to 
x
\%) in 2020, a 
x
\% (
x
\% to 
x
\%) increase. 
In contrast, the lowest increase in MC coverage in these countries was in 
x
, from 
x
\% (
x
\% to 
x
\%) in 2006 to 
x
\% (
x
\% to 
x
\%) in 2020, a 
x
\% (
x
\% to 
x
\%) increase.

However, these figures belie significant subnational heterogeneity in MC coverage amongst VMMC prioirty countries.

x
/
x
 districts in these countries were e
x
pected to have reached 90\% MC coverage amongst 10-29 year olds in 2020, up from 
x
/
x
 in 2006, before VMMC programmes began. 
Even in 
x
, where MC was projected to be lowest in 2020, 
x
/
x
 districts were forecasted to have reached 90\% MC coverage for men aged 10-29 years old. 
MC coverage there increased from 
x
(
x
to 
x
) in 2006 to 
x
(
x
to 
x
) in 2020, a 
x
(
x
to 
x) increase.