### 2024-04-12 paper/paper_cals.Rmd
Finish section 4

ub
title: "Circumcision Paper Text"
author: "Paddy O'Toole"
date: "2023-11-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- TODO: Change year from 2020 to 2023 -->

```{r, echo = FALSE}
# libraries
library(dplyr, warn.conflicts = FALSE)

# Metadata 
spec_agegroup <- "15-29"
spec_years <- c(2006, 2020)

ssa_iso3 <- sort(c(
  "AGO", "BDI", "BEN", "BFA", "BWA", "CAF", "CIV", "CMR", "COD",
  "COG", "ETH", "GAB", "GHA", "GIN", "GMB", "GNB", "GNQ", "KEN",
  "LBR", "LSO", "MLI", "MOZ", "MWI", "NAM", "NER", "NGA", "RWA",
  "SEN", "SLE", "SWZ", "TCD", "TGO", "TZA", "UGA", "ZAF", "ZMB", "ZWE"
))
```

```{r, echo = FALSE}
# Functions

# change numbers to be expressed in "millions"
to_mill <- \(x) round(x / 1e6, 3)

# change decimal to percentage
to_percentage <- \(x) paste0(round(x * 100, 2), "%")

# translate iso3 to country name
to_cntry_name <- function(x) {
  x_cntry = countrycode::countrycode(x, "iso3c", "country.name")
  # special cases
  case_when(
      x_cntry == "Congo - Brazzaville" ~ "Congo",
      x_cntry == "Congo - Kinshasa"    ~ "DR Congo",
      TRUE                             ~ x_cntry
  )
}

# simplify type for faster filtering
simplify_type <- function(dat) {
  dat %>% 
    mutate(type = case_when(
      grepl("MMC", type) ~ "MMC", 
      grepl("TMC", type) ~ "TMC", 
      TRUE               ~ "MC"
    ))
}

# tidy up aggregated age data
tidy_age <- function(dat, sig_fig = 2) {
  dat %>% 
    mutate(across(contains("average_age"), \(x) round(x, sig_fig))) %>% 
    rename(
      mean = average_age, lower = average_age_lower, upper = average_age_upper
    ) %>% 
    simplify_type()
}
```

# Results

## Subnational Variation in Total, Medical \& Traditional Circumcision over time

<!-- What else could we add to this section? Section 2 is very large!!!
- Could talk about non-VMMC vs VMMC more anyway -->

```{r, echo = FALSE}
x1 <- readRDS("data/inlines/01_inlines.RDS")
for (i in seq_along(x1)) {
  assign(names(x1)[i], x1[[i]])
}
iso3_lowest_highest_cov <- simplify_type(iso3_lowest_highest_cov)
iso3_cov_diffs <- simplify_type(iso3_cov_diffs)
```

<!-- I surely need to add something about 10-29/15-29 year olds here? -->
<!-- Plot is of 15-29 year olds, but then I talk only about whole pop here -->
In 2006, the year before VMMC programme implementation began, an estimated 
`r n_circs[[1]] %>% filter(year == 2006, type == "MC") %>% pull(mean) %>% to_mill()` 
million men (95\% CI: 
`r n_circs[[1]] %>% filter(year == 2006, type == "MC") %>% pull(lower) %>% to_mill()` 
-
`r n_circs[[1]] %>% filter(year == 2006, type == "MC") %>% pull(upper) %>% to_mill()` 
million) were circumcised in the 33 countries in SSA for which we have valid 
circumcision data. 
Of these, 
`r n_circs[[1]] %>% filter(year == 2006, type == "MMC") %>% pull(mean) %>% to_mill()` 
million 
(`r n_circs[[1]] %>% filter(year == 2006, type == "MMC") %>% pull(lower) %>% to_mill()` 
- 
`r n_circs[[1]] %>% filter(year == 2006, type == "MMC") %>% pull(upper) %>% to_mill()` 
million) were MMCS, and 
`r n_circs[[1]] %>% filter(year == 2006, type == "TMC") %>% pull(mean) %>% to_mill()` 
million 
(`r n_circs[[1]] %>% filter(year == 2006, type == "TMC") %>% pull(lower) %>% to_mill()` 
- 
`r n_circs[[1]] %>% filter(year == 2006, type == "TMC") %>% pull(upper) %>% to_mill()` 
million) were TMCS. 
This contrasts with 2020, where an estimated 
`r n_circs[[1]] %>% filter(year == 2020, type == "MC") %>% pull(mean) %>% to_mill()` 
million males 
(`r n_circs[[1]] %>% filter(year == 2020, type == "MC") %>% pull(lower) %>% to_mill()` 
- 
`r n_circs[[1]] %>% filter(year == 2020, type == "MC") %>% pull(upper) %>% to_mill()` 
million) were circumcised in these same countries, of which 
`r n_circs[[1]] %>% filter(year == 2020, type == "MMC") %>% pull(mean) %>% to_mill()` 
million 
(`r n_circs[[1]] %>% filter(year == 2020, type == "MMC") %>% pull(lower) %>% to_mill()` 
- 
`r n_circs[[1]] %>% filter(year == 2020, type == "MMC") %>% pull(upper) %>% to_mill()` 
million) were performed as MMCS, and 
`r n_circs[[1]] %>% filter(year == 2020, type == "TMC") %>% pull(mean) %>% to_mill()` 
million 
(`r n_circs[[1]] %>% filter(year == 2020, type == "TMC") %>% pull(lower) %>% to_mill()` 
- 
`r n_circs[[1]] %>% filter(year == 2020, type == "TMC") %>% pull(upper) %>% to_mill()` 
million) as TMCS. 
This constituted an increase of 
`r n_circs[[2]] %>% filter(type == "MC") %>% pull(mean) %>% to_mill()` 
million 
(`r n_circs[[2]] %>% filter(type == "MC") %>% pull(lower) %>% to_mill()` 
- 
`r n_circs[[2]] %>% filter(type == "MC") %>% pull(upper) %>% to_mill()` 
million) newly circumcised males between 2006 and 2020, with 
`r n_circs[[2]] %>% filter(type == "MMC") %>% pull(mean) %>% to_mill()` 
million 
(`r n_circs[[2]] %>% filter(type == "MMC") %>% pull(lower) %>% to_mill()` 
- 
`r n_circs[[2]] %>% filter(type == "MMC") %>% pull(upper) %>% to_mill()` 
million) new MMCs performed, and 
`r n_circs[[2]] %>% filter(type == "TMC") %>% pull(mean) %>% to_mill()` 
million 
(`r n_circs[[2]] %>% filter(type == "TMC") %>% pull(lower) %>% to_mill()` 
- 
`r n_circs[[2]] %>% filter(type == "TMC") %>% pull(upper) %>% to_mill()` 
million) new TMCs performed. 
The disparity between the number of MCs performed and the number of MMCs performed can be partly explained by recalling that for five countries, Liberia, Senegal, Niger, Guinea and the Democratic Republic of the Congo, we have no circumcision type information, and so can only estimate MC, not MMC or TMC.
This translated to a change in mean MC coverage for the SSA region from 
`r covs[[1]] %>% filter(year == 2006, type == "MC") %>% pull(mean) %>% to_percentage()` 
(`r covs[[1]] %>% filter(year == 2006, type == "MC") %>% pull(upper) %>% to_percentage()` 
- 
`r covs[[1]] %>% filter(year == 2006, type == "MC") %>% pull(lower) %>% to_percentage()` ) 
in 2006 to 
`r covs[[1]] %>% filter(year == 2020, type == "MC") %>% pull(mean) %>% to_percentage()` 
(`r covs[[1]] %>% filter(year == 2020, type == "MC") %>% pull(upper) %>% to_percentage()` 
- 
`r covs[[1]] %>% filter(year == 2020, type == "MC") %>% pull(lower) %>% to_percentage()`)
in 2020, or an increase of
`r covs[[2]] %>% filter(type == "MC") %>% pull(mean) %>% to_percentage()` 
(`r covs[[2]] %>% filter(type == "MC") %>% pull(lower) %>% to_percentage()` 
- 
`r covs[[2]] %>% filter(type == "MC") %>% pull(upper) %>% to_percentage()`).
MMC coverage increased from 
`r covs[[1]] %>% filter(year == 2006, type == "MMC") %>% pull(mean) %>% to_percentage()` 
(`r covs[[1]] %>% filter(year == 2006, type == "MMC") %>% pull(lower) %>% to_percentage()` 
- 
`r covs[[1]] %>% filter(year == 2006, type == "MMC") %>% pull(upper) %>% to_percentage()` ) 
in 2006 to 
`r covs[[1]] %>% filter(year == 2020, type == "MMC") %>% pull(mean) %>% to_percentage()` 
(`r covs[[1]] %>% filter(year == 2020, type == "MMC") %>% pull(lower) %>% to_percentage()` 
- 
`r covs[[1]] %>% filter(year == 2020, type == "MMC") %>% pull(upper) %>% to_percentage()`) 
in 2020, an increase of 
`r covs[[2]] %>% filter(type == "MMC") %>% pull(mean) %>% to_percentage()` 
(`r covs[[2]] %>% filter(type == "MMC") %>% pull(lower) %>% to_percentage()` 
- 
`r covs[[2]] %>% filter(type == "MMC") %>% pull(upper) %>% to_percentage()`).
while TMC changed from 
`r covs[[1]] %>% filter(year == 2006, type == "TMC") %>% pull(mean) %>% to_percentage()` 
(`r covs[[1]] %>% filter(year == 2006, type == "TMC") %>% pull(lower) %>% to_percentage()` 
- 
`r covs[[1]] %>% filter(year == 2006, type == "TMC") %>% pull(upper) %>% to_percentage()`)
in 2006 to 
`r covs[[1]] %>% filter(year == 2020, type == "TMC") %>% pull(mean) %>% to_percentage()` 
(`r covs[[1]] %>% filter(year == 2020, type == "TMC") %>% pull(lower) %>% to_percentage()` 
- 
`r covs[[1]] %>% filter(year == 2020, type == "TMC") %>% pull(upper) %>% to_percentage()`)
in 2020, a decrease of
`r covs[[2]] %>% filter(type == "TMC") %>% mutate(mean = mean * -1) %>% pull(mean) %>% to_percentage()` 
(`r covs[[2]] %>% filter(type == "TMC") %>% mutate(upper = upper * -1) %>% pull(upper) %>% to_percentage()` 
- 
`r covs[[2]] %>% filter(type == "TMC") %>%  mutate(lower = lower * -1) %>%  pull(lower) %>% to_percentage()`).

<!-- TODO: Need to change this to be filtered by year as well!! -->
MC coverage in each country in 2006 ranged from 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2006) %>% slice(2) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2006) %>% slice(2) %>% pull(lower) %>% to_percentage()`
- 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2006) %>% slice(2) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2006) %>% slice(2) %>% pull(iso3) %>% to_cntry_name()`
to 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2006) %>% slice(1) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2006) %>% slice(1) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2006) %>% slice(1) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2006) %>% slice(1) %>% pull(iso3) %>% to_cntry_name()`, 
while MMC in 2006 ranged from 
`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2006) %>% slice(2) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2006) %>% slice(2) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2006) %>% slice(2) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2006) %>% slice(2) %>% pull(iso3) %>% to_cntry_name()`
to
`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2006) %>% slice(1) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2006) %>% slice(1) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2006) %>% slice(1) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2006) %>% slice(1) %>% pull(iso3) %>% to_cntry_name()`
and TMC ranged from 
`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2006) %>% slice(2) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2006) %>% slice(2) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2006) %>% slice(2) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2006) %>% slice(2) %>% pull(iso3) %>% to_cntry_name()`
to 
`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2006) %>% slice(1) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2006) %>% slice(1) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2006) %>% slice(1) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2006) %>% slice(1) %>% pull(iso3) %>% to_cntry_name()`.
In contrast, in 2020 MC coverage was estimated to have ranged from 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2020) %>% slice(2) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2020) %>% slice(2) %>% pull(lower) %>% to_percentage()`
- 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2020) %>% slice(2) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2020) %>% slice(2) %>% pull(iso3) %>% to_cntry_name()`
to 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2020) %>% slice(1) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2020) %>% slice(1) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2020) %>% slice(1) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "MC", year == 2020) %>% slice(1) %>% pull(iso3) %>% to_cntry_name()`, 
while MMC ranged from 
`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2020) %>% slice(2) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2020) %>% slice(2) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2020) %>% slice(2) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2020) %>% slice(2) %>% pull(iso3) %>% to_cntry_name()`
to
`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2020) %>% slice(1) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2020) %>% slice(1) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2020) %>% slice(1) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "MMC", year == 2020) %>% slice(1) %>% pull(iso3) %>% to_cntry_name()`
and TMC ranged from 
`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2020) %>% slice(2) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2020) %>% slice(2) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2020) %>% slice(2) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2020) %>% slice(2) %>% pull(iso3) %>% to_cntry_name()`
to 
`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2020) %>% slice(1) %>% pull(mean) %>% to_percentage()`
(`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2020) %>% slice(1) %>% pull(lower) %>% to_percentage()`
- 
(`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2020) %>% slice(1) %>% pull(upper) %>% to_percentage()`)
in 
`r iso3_lowest_highest_cov %>% filter(type == "TMC", year == 2020) %>% slice(1) %>% pull(iso3) %>% to_cntry_name()`.
<!-- Anything else to be said about this? -->

The largest percentage increase in MC coverage from 2006 to 2020 was 
in 
`r iso3_cov_diffs %>% filter(type == "MC") %>% slice(1) %>% pull(iso3) %>% to_cntry_name()`
at
`r iso3_cov_diffs %>% filter(type == "MC") %>% slice(1) %>% pull(mean) %>% to_percentage()`
(`r iso3_cov_diffs %>% filter(type == "MC") %>% slice(1) %>% pull(lower) %>% to_percentage()`
-
`r iso3_cov_diffs %>% filter(type == "MC") %>% slice(1) %>% pull(upper) %>% to_percentage()`),
from 
`r iso3_cov_2006_2020 %>% filter(type == "MC", year == 2006) %>% pull(mean) %>% to_percentage()`
(`r iso3_cov_2006_2020 %>% filter(type == "MC", year == 2006) %>% pull(lower) %>% to_percentage()`
-
`r iso3_cov_2006_2020 %>% filter(type == "MC", year == 2006) %>% pull(upper) %>% to_percentage()`)
in 2006 to 
`r iso3_cov_2006_2020 %>% filter(type == "MC", year == 2020) %>% pull(mean) %>% to_percentage()`
(`r iso3_cov_2006_2020 %>% filter(type == "MC", year == 2020) %>% pull(lower) %>% to_percentage()`
-
`r iso3_cov_2006_2020 %>% filter(type == "MC", year == 2020) %>% pull(upper) %>% to_percentage()`)
in 2020,
while the highest percentage increase in MMC was also in 
`r iso3_cov_diffs %>% filter(type == "MMC") %>% slice(1) %>% pull(iso3) %>% to_cntry_name()`
at
`r iso3_cov_diffs %>% filter(type == "MMC") %>% slice(1) %>% pull(mean) %>% to_percentage()`
(`r iso3_cov_diffs %>% filter(type == "MMC") %>% slice(1) %>% pull(lower) %>% to_percentage()`
-
`r iso3_cov_diffs %>% filter(type == "MMC") %>% slice(1) %>% pull(upper) %>% to_percentage()`),
from 
`r iso3_cov_2006_2020 %>% filter(type == "MMC", year == 2006) %>% pull(mean) %>% to_percentage()`
(`r iso3_cov_2006_2020 %>% filter(type == "MMC", year == 2006) %>% pull(lower) %>% to_percentage()`
-
`r iso3_cov_2006_2020 %>% filter(type == "MMC", year == 2006) %>% pull(upper) %>% to_percentage()`)
in 2006 to 
`r iso3_cov_2006_2020 %>% filter(type == "MMC", year == 2020) %>% pull(mean) %>% to_percentage()`
(`r iso3_cov_2006_2020 %>% filter(type == "MMC", year == 2020) %>% pull(lower) %>% to_percentage()`
-
`r iso3_cov_2006_2020 %>% filter(type == "MMC", year == 2020) %>% pull(upper) %>% to_percentage()`)
in 2020. 
For TMC, the largest percentage decrease between 2006 and 2020 was in 
`r iso3_cov_diffs %>% filter(type == "TMC") %>% slice(1) %>% pull(iso3) %>% to_cntry_name()`
at
`r iso3_cov_diffs %>% filter(type == "TMC") %>% slice(1) %>% pull(mean) %>% to_percentage()`
(`r iso3_cov_diffs %>% filter(type == "TMC") %>% slice(1) %>% pull(upper) %>% to_percentage()`
-
`r iso3_cov_diffs %>% filter(type == "TMC") %>% slice(1) %>% pull(lower) %>% to_percentage()`),
from 
`r iso3_cov_2006_2020 %>% filter(type == "TMC", year == 2006) %>% pull(mean) %>% to_percentage()`
(`r iso3_cov_2006_2020 %>% filter(type == "TMC", year == 2006) %>% pull(lower) %>% to_percentage()`
-
`r iso3_cov_2006_2020 %>% filter(type == "TMC", year == 2006) %>% pull(upper) %>% to_percentage()`)
in 2006 to 
`r iso3_cov_2006_2020 %>% filter(type == "TMC", year == 2020) %>% pull(mean) %>% to_percentage()`
(`r iso3_cov_2006_2020 %>% filter(type == "TMC", year == 2020) %>% pull(lower) %>% to_percentage()`
-
`r iso3_cov_2006_2020 %>% filter(type == "TMC", year == 2020) %>% pull(upper) %>% to_percentage()`)
in 2020.

The number of annual MCs performed increased from 
`r non_cum_n_performed %>% filter(type == "MC", year == 2006) %>% pull(mean) %>% to_mill()` 
million
(`r non_cum_n_performed %>% filter(type == "MC", year == 2006) %>% pull(lower) %>% to_mill()` 
-
`r non_cum_n_performed %>% filter(type == "MC", year == 2006) %>% pull(upper) %>% to_mill()` 
million) in 2006 to 
`r non_cum_n_performed %>% filter(type == "MC", year == 2006) %>% pull(mean) %>% to_mill()` 
million 
(`r non_cum_n_performed %>% filter(type == "MC", year == 2006) %>% pull(lower) %>% to_mill()` 
-
`r non_cum_n_performed %>% filter(type == "MC", year == 2006) %>% pull(upper) %>% to_mill()` 
million) in 2020, a growth of 
`r non_cum_n_performed_diff %>% filter(type == "MC") %>% pull(mean) %>% to_mill()` 
million (
`r non_cum_n_performed_diff %>% filter(type == "MC") %>% pull(lower) %>% to_mill()`
-
`r non_cum_n_performed_diff %>% filter(type == "MC") %>% pull(upper) %>% to_mill()` 
million) annually, averaged across all interceding years. 
Amongst these, 
`r non_cum_n_performed %>% filter(type == "MC", year == 2006) %>% pull(mean) %>% to_mill()` 
million
(`r non_cum_n_performed %>% filter(type == "MC", year == 2006) %>% pull(lower) %>% to_mill()` 
-
`r non_cum_n_performed %>% filter(type == "MC", year == 2006) %>% pull(upper) %>% to_mill()` 
million) were MMCs in 2006, while in 2020 
`r non_cum_n_performed %>% filter(type == "MC", year == 2020) %>% pull(mean) %>% to_mill()` 
million
(`r non_cum_n_performed %>% filter(type == "MC", year == 2020) %>% pull(lower) %>% to_mill()` 
-
`r non_cum_n_performed %>% filter(type == "MC", year == 2020) %>% pull(upper) %>% to_mill()` 
million) MMCs were performed, representing an increase in 
million 
(`r non_cum_n_performed_diff %>% filter(type == "MMC") %>% pull(lower) %>% to_mill()` 
-
`r non_cum_n_performed_diff %>% filter(type == "MMC") %>% pull(upper) %>% to_mill()` 
million) in MMCs performed annually. 
The number of TMCs performed annually did increase slightly from 2006, from 
`r non_cum_n_performed %>% filter(type == "TMC", year == 2006) %>% pull(mean) %>% to_mill()` 
million 
(
`r non_cum_n_performed %>% filter(type == "TMC", year == 2006) %>% pull(lower) %>% to_mill()` 
- 
`r non_cum_n_performed %>% filter(type == "TMC", year == 2006) %>% pull(upper) %>% to_mill()` 
million) to 
`r non_cum_n_performed %>% filter(type == "TMC", year == 2020) %>% pull(mean) %>% to_mill()`
million
(
`r non_cum_n_performed %>% filter(type == "TMC", year == 2020) %>% pull(lower) %>% to_mill()`
- 
`r non_cum_n_performed %>% filter(type == "TMC", year == 2020) %>% pull(upper) %>% to_mill()`
million) in 2020. However, 
TMC coverage actually decreased in many countries from 2006 to 2020. 
<!-- Should I give a number here? -->
This is the focus of section \ref{sec:org5fb5e18}.

<!-- *Second Plot* -->
<!-- Want to show here how ??? -->
Within countries, the median difference in MC coverage between districts with lowest and highest coverage in 2020 was `r district_diffs_vmmc$median_cov_diffs %>% filter(type == "MC") %>% pull(mean) %>% to_percentage()`
The smallest variation was in
`r district_diffs_vmmc$country_cov_diffs %>% filter(type == "MC") %>% pull(iso3) %>% last() %>% to_cntry_name()`, 
at 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MC") %>% slice(2) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MC") %>% slice(2) %>% pull(area_name)`
to
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MC") %>% slice(1) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MC") %>% slice(1) %>% pull(area_name)`
,
while the largest variation was in 
`r district_diffs_vmmc$country_cov_diffs %>% filter(type == "MC") %>% pull(iso3) %>% first() %>% to_cntry_name()`, 
at 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MC") %>% slice(4) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MC") %>% slice(4) %>% pull(area_name)`
to
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MC") %>% slice(3) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MC") %>% slice(3) %>% pull(area_name)`.
The median difference within countries in MMC coverage between districts with lowest and highest coverage in 2020 was `r district_diffs_vmmc$median_cov_diffs %>% filter(type == "MMC") %>% pull(mean) %>% to_percentage()`
The smallest variation was in
`r district_diffs_vmmc$country_cov_diffs %>% filter(type == "MMC") %>% pull(iso3) %>% last() %>% to_cntry_name()`, 
at
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MMC") %>% slice(2) %>% pull(mean) %>% to_percentage()`
\%
in
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MMC") %>% slice(2) %>% pull(area_name)`
to
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MMC") %>% slice(1) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MMC") %>% slice(1) %>% pull(area_name)`,
while the largest variation was in 
`r district_diffs_vmmc$country_cov_diffs %>% filter(type == "MMC") %>% pull(iso3) %>% first() %>% to_cntry_name()`, 
at 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MMC") %>% slice(4) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MMC") %>% slice(4) %>% pull(area_name)`
to
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MMC") %>% slice(3) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "MMC") %>% slice(3) %>% pull(area_name)`.
For TMC coverage, the median difference between districts in 2020 was 
`r district_diffs_vmmc$median_cov_diffs %>% filter(type == "TMC") %>% pull(mean) %>% to_percentage()`
The smallest variation was in
`r district_diffs_vmmc$country_cov_diffs %>% filter(type == "TMC") %>% pull(iso3) %>% last() %>% to_cntry_name()`, 
at
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "TMC") %>% slice(2) %>% pull(mean) %>% to_percentage()`
\%
in
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "TMC") %>% slice(2) %>% pull(area_name)`
to
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "TMC") %>% slice(1) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "TMC") %>% slice(1) %>% pull(area_name)`,
while the largest variation was in 
`r district_diffs_vmmc$country_cov_diffs %>% filter(type == "TMC") %>% pull(iso3) %>% first() %>% to_cntry_name()`, 
at 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "TMC") %>% slice(4) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "TMC") %>% slice(4) %>% pull(area_name)`
to
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "TMC") %>% slice(3) %>% pull(mean) %>% to_percentage()`
\%
in 
`r district_diffs_vmmc$country_cov_highest_lowest %>% filter(type == "TMC") %>% slice(3) %>% pull(area_name)`.
<!-- Anything else? --->

## Subnational Variation in total, medical \& traditional circumcision across different ages

```{r, echo = FALSE}
x2 <- readRDS("data/inlines/02_inlines.RDS")
for (i in seq_along(x2)) {
  assign(names(x2)[i], x2[[i]])
}
age_group_cov <- sapply(
  age_group_cov, simplify_type, simplify = FALSE, USE.NAMES = TRUE
)
age_group_cov_non_vmmc <- sapply(
  age_group_cov_non_vmmc, simplify_type, simplify = FALSE, USE.NAMES = TRUE
)
age_group_cov_vmmc <- sapply(
  age_group_cov_vmmc, simplify_type, simplify = FALSE, USE.NAMES = TRUE
)
```

<!-- TODO: Replace WCA/ECA with VMMC/Non-VMMC -->

Age at circumcision differed greatly by circumcision type and location. 
The five-year age group in 2006, the year before VMMC programme implementation began, 
which had the highest MC coverage across the SSA region was
`r age_group_cov$fyag_highest_covs %>% filter(type == "MC", year == 2006) %>% pull(age_group)`, 
at 
`r age_group_cov$fyag_highest_covs %>% filter(type == "MC", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$fyag_highest_covs %>% filter(type == "MC", year == 2006) %>% pull(lower) %>% to_percentage()`
-
`r age_group_cov$fyag_highest_covs %>% filter(type == "MC", year == 2006) %>% pull(upper) %>% to_percentage()`
)
. 
In contrast, the age group with the highest MC coverage in 2020 was 
`r age_group_cov$fyag_highest_covs %>% filter(type == "MC", year == 2020) %>% pull(age_group)`,
at
`r age_group_cov$fyag_highest_covs %>% filter(type == "MC", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$fyag_highest_covs %>% filter(type == "MC", year == 2020) %>% pull(lower) %>% to_percentage()`
-
`r age_group_cov$fyag_highest_covs %>% filter(type == "MC", year == 2020) %>% pull(upper) %>% to_percentage()`
).
The five-year age group in 2006 which had the highest MMC coverage was
`r age_group_cov$fyag_highest_covs %>% filter(type == "MMC", year == 2006) %>% pull(age_group)`, 
at 
`r age_group_cov$fyag_highest_covs %>% filter(type == "MMC", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$fyag_highest_covs %>% filter(type == "MMC", year == 2006) %>% pull(lower) %>% to_percentage()`
-
`r age_group_cov$fyag_highest_covs %>% filter(type == "MMC", year == 2006) %>% pull(upper) %>% to_percentage()`
)
. 
The age group with the highest MMC coverage in 2020 was also
`r age_group_cov$fyag_highest_covs %>% filter(type == "MMC", year == 2020) %>% pull(age_group)`,
, now at
`r age_group_cov$fyag_highest_covs %>% filter(type == "MMC", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$fyag_highest_covs %>% filter(type == "MMC", year == 2020) %>% pull(lower) %>% to_percentage()`
-
`r age_group_cov$fyag_highest_covs %>% filter(type == "MMC", year == 2020) %>% pull(upper) %>% to_percentage()`
).
The five-year age group in 2006 which had the highest TMC coverage was
`r age_group_cov$fyag_highest_covs %>% filter(type == "TMC", year == 2006) %>% pull(age_group)`, 
at 
`r age_group_cov$fyag_highest_covs %>% filter(type == "TMC", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$fyag_highest_covs %>% filter(type == "TMC", year == 2006) %>% pull(lower) %>% to_percentage()`
-
`r age_group_cov$fyag_highest_covs %>% filter(type == "TMC", year == 2006) %>% pull(upper) %>% to_percentage()`
)
. 
The age group with the highest TMC coverage in 2020 was
`r age_group_cov$fyag_highest_covs %>% filter(type == "TMC", year == 2020) %>% pull(age_group)`,
,
`r age_group_cov$fyag_highest_covs %>% filter(type == "TMC", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$fyag_highest_covs %>% filter(type == "TMC", year == 2020) %>% pull(lower) %>% to_percentage()`
-
`r age_group_cov$fyag_highest_covs %>% filter(type == "TMC", year == 2020) %>% pull(upper) %>% to_percentage()`
).
This decrease in the five-year age group with the highest MC coverage, increase in the 
age group for TMC with the highest coverage (as well as a decrease in coverage for 
the highest five-year age group), and increase in overall MMC coverage across the SSA region is 
indicative of the effect of VMMC programme implementation on the region.

<!-- TODO Still: 
- 15-29 year olds increase in MC/TMC/MMC (already written about?)
- Contrast to 30-49 year olds somehow, showing different patterns of circ
(increasing more for MC/MMC, decreasing more for TMC)
- Compare differences in VMMC and non-VMMC, in terms of the above
-->
We may compare changes in circumcision amongst two age groups; 15-29 year olds, which 
are a particular focus for VMMC implementation, and 30-49 year olds, which are not. 
MC coverage amongst 15-29 year olds increased from 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020. 
In contrast, MC coverage amongst 30-49 year olds only increased from 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020. 
Therefore, MC coverage across SSA saw greater increases in this younger age group than in this older one, in line 
with VMMC programme age focus. 
Similarly, MMC coverage amongst 15-29 year olds increased from 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020, while MMC coverage amongst 30-49 year olds increased from 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020. 
Therefore, MMC coverage across SSA also saw greater increases amongst 15-29 year olds 
than amongst 30-49 year olds. 
The changes seen in TMC, however, contrast to these larger increases in MC and MMC coverage for 15-29 year olds. 
TMC coverage amongst 15-29 year olds decreased from 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020, while TMC coverage amongst 30-49 year olds decreased from 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020. 
Thus, TMC coverage decreased less in absolute terms amongst 30-49 year olds than amongst 
15-29 year olds, which is line with the greater focus on performing MMCs amongst this 
younger age group.

Finally, we can contrast patterns in circumcision pattern for different age groups
amongst VMMC and non-VMMC priority countries, where differences in 2006 and 2020 
levels and patterns of circumcision seem to be more pronounced.

<!-- For VMMC countries, five year age group MC, MMC, TMC -->
For VMMC priority countries, five-year age group MC coverage 2006 was highest amongst 
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MC") %>% pull(age_group)`
year olds, at 
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MC") %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MC") %>% pull(lower) %>% to_percentage()`
 - 
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MC") %>% pull(upper) %>% to_percentage()`
), but in 2020 was highest amongst
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MC") %>% pull(age_group)`
year olds, at
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MC") %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MC") %>% pull(lower) %>% to_percentage()`
 - 
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MC") %>% pull(upper) %>% to_percentage()`
), respectively. 
This highlights both the increased prevalence of MC in VMMC priority countries, and the
downward shift in age profile of those who are circumcised, since VMMC programme 
implementation began. 
In 2006 MMC coverage was highest amongst 
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MMC") %>% pull(age_group)`
 year olds, at 
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MMC") %>% pull(mean) %>% to_percentage()`
 (
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MMC") %>% pull(lower) %>% to_percentage()`
 - 
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MMC") %>% pull(upper) %>% to_percentage()`
). 
This contrasts with 2020, where MMC coverage was highest amongst 
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MMC") %>% pull(age_group)`
 year olds, at
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MMC") %>% pull(mean) %>% to_percentage()`
 (
x
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MMC") %>% pull(lower) %>% to_percentage()`
 - 
x
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MMC") %>% pull(upper) %>% to_percentage()`
).
This shows that much of the downward shift in age profile of circumcised individuals since 2006 
has been "driven" by an increase in the number of MMCs performed amongst younger individuals. 
In 2006, TMC coverage was greatest in 
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(age_group)`
year olds, at 
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(mean) %>% to_percentage()`
 (
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(lower) %>% to_percentage()`
 - 
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(upper) %>% to_percentage()`
). while in 2020 it was most prevalent in
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(age_group)`
year olds, at 
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(mean) %>% to_percentage()`
 (
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(lower) %>% to_percentage()`
 - 
`r age_group_cov_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(upper) %>% to_percentage()`.
This perhaps highlights that VMMC implementation has been focused on areas where 
TMC has not historically been widely prevalent, and so seems to have had little effect on patterns of TMC in 
VMMC priority countries, both in terms of its age profile and prevalence.

We can constrast this for the non-VMMC priority countries in SSA.
The five-year age group MC coverage in both 2006 and 2020 was highest amongst 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MC") %>% pull(age_group)`
year olds, at 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MC") %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MC") %>% pull(lower) %>% to_percentage()`
 - 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MC") %>% pull(upper) %>% to_percentage()`
) and 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MC") %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MC") %>% pull(lower) %>% to_percentage()`
 - 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MC") %>% pull(upper) %>% to_percentage()`
), respectively. 
This indicates that MC has increased slightly in non-VMMC priority countries, but the age profile (in terms of the five-year age group with the highest coverage) hasn't noticably changed. 
However, in 2006 MMC coverage was highest amongst 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MMC") %>% pull(age_group)`
 year olds, at 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MMC") %>% pull(mean) %>% to_percentage()`
 (
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MMC") %>% pull(lower) %>% to_percentage()`
 - 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "MMC") %>% pull(upper) %>% to_percentage()`
). 
This contrasts with 2020, where MMC coverage was highest amongst 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MMC") %>% pull(age_group)`
 year olds, at
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MMC") %>% pull(mean) %>% to_percentage()`
 (
x
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MMC") %>% pull(lower) %>% to_percentage()`
 - 
x
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2020, type == "MMC") %>% pull(upper) %>% to_percentage()`
), showing an interesting shift upwards in age for MMCs in non-VMMC priority countries, as a result of VMMC focus on young adults, as well as a general increase in MMC coverage in these countries. 
In 2006, TMC coverage was greatest in 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(age_group)`
year olds, at 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(mean) %>% to_percentage()`
 (
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(lower) %>% to_percentage()`
 - 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(upper) %>% to_percentage()`
). while in 2020 it was most prevalent in
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(age_group)`
year olds, at 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(mean) %>% to_percentage()`
 (
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(lower) %>% to_percentage()`
 - 
`r age_group_cov_non_vmmc$fyag_highest_covs %>% filter(year == 2006, type == "TMC") %>% pull(upper) %>% to_percentage()`, highlighting a shift upwards in TMCs in VMMC priority countries, 
likely resulting from previously TMCs being performed as TMCs.

Similarly, we can compare the 2006 and 2020 circumcision levels amongst 15-29 and 30-49 year olds
in both VMMC and non-VMMC countries, respectively.

In VMMC priority countries, MC coverage amongst 15-29 year olds increased from 
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020. 
In contrast, MC coverage in 30-49 year olds went from 
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020. 
Therefore, MC coverage in VMMC priority countries increased amongst 15-29 year olds at a higher rate than amongst 30-49 year olds, and indeed greater than in SSA as a whole, highlighting the effects of VMMC implementation in these countries since 2006. 
Similarly, MMC coverage amongst 15-29 year olds increased from 
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020, while MMC coverage amongst 30-49 year olds increased from 
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020. 
Therefore, MMC coverage in VMMC priority countries also saw greater increases amongst 15-29 year olds 
than amongst 30-49 year olds, and greater increases than for SSA as a whole. 
Again, TMC coverage amongst both fifteen year age groups constrast to these increases for MC and MMC. 
TMC coverage amongst 15-29 year olds decreased from 
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020, while TMC coverage amongst 30-49 year olds decreased from 
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020. 
This decrease in TMC coverage can partly be attributed to a combination of increased population in non-TMC practising areas, but TMC coverage also decreased less in absolute terms amongst 30-49 year olds than amongst 
15-29 year olds, which is line with the greater focus on performing MMCs, and hence replacing TMCs, amongst this 
younger age group.

In contrast, for non-VMMC priority countries, MC coverage in 15-29 year olds increased from 
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "15-29", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020, staying relatively similar. 
MC coverage in 30-49 year olds also increased from 
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MC", age_group == "30-49", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020, so again, when taking our credible intervals into account, there was little 
increase amongst this age group in non-VMMC priority countries. 
It is interesting that MC seems to have increased more amongst 30-49 year olds 
than amongst 15-29, defying the trend amongst VMMC priority countries and SSA as a whole. 
MMC coverage amongst 15-29 year olds increased from 
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "15-29", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020, while MMC coverage amongst 30-49 year olds increased from 
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "MMC", age_group == "30-49", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020. 
Hence, MMC coverage in non-VMMC countries also saw greater increases amongst 30-49 year olds 
than in 15-29 year olds, but not much change overall when factoring in our uncertainty.
TMC coverage amongst 15-29 year olds decreased from 
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "15-29", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020, while TMC coverage amongst 30-49 year olds decreased from 
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2006) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2006) %>% pull(upper) %>% to_percentage()`
) in 2006
to
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2020) %>% pull(lower) %>% to_percentage()`
- 
`r age_group_cov_non_vmmc$spec_agegroups_mean_cov %>% filter(type == "TMC", age_group == "30-49", year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020. 
Thus, TMC coverage decreased less in absolute terms amongst 30-49 year olds than amongst 
15-29 year olds, which is line with the greater focus on performing MMCs amongst this 
younger age group. 
TMC coverage decreased more amongst non-VMMC countries for both age groups, suggesting 
likely replacement of historical TMC with MMC.

## Subnational Heterogeneity in Age at Medical and Traditional Circumcision

```{r}
x3 <- readRDS("data/inlines/03_inlines.RDS")
for (i in seq_along(x3)) {
  assign(names(x3)[i], x3[[i]])
}
aggr_mean_circ_age <- tidy_age(aggr_mean_circ_age)
aggr_mean_circ_age_non_vmmc <- tidy_age(aggr_mean_circ_age_non_vmmc)
aggr_mean_circ_age_vmmc <- tidy_age(aggr_mean_circ_age_vmmc)
```

Fill in! 
- Talk mainly about average age at which MCs/MMCs/TMCs performed in 2006 and 2020
- Also compare VMMC and non-VMMC

The mean age at circumcision for MC in 2006 was 
`r aggr_mean_circ_age %>% filter(year == 2006, type == "MC") %>% pull(mean)`
(
`r aggr_mean_circ_age %>% filter(year == 2006, type == "MC") %>% pull(lower)`
- 
`r aggr_mean_circ_age %>% filter(year == 2006, type == "MC") %>% pull(upper)`
), while in 2020 it was
`r aggr_mean_circ_age %>% filter(year == 2020, type == "MC") %>% pull(mean)`
(
`r aggr_mean_circ_age %>% filter(year == 2020, type == "MC") %>% pull(lower)`
- 
`r aggr_mean_circ_age %>% filter(year == 2020, type == "MC") %>% pull(upper)`
).
<!-- Therefore, factoring in uncertainty there doesn't seem to have been a 
significant drop in mean age at MC from 2006 to 2020. -->
Similarly, the mean age at circumcision for MMC in 2006 was 
`r aggr_mean_circ_age %>% filter(year == 2006, type == "MMC") %>% pull(mean)`
(
`r aggr_mean_circ_age %>% filter(year == 2006, type == "MMC") %>% pull(lower)`
- 
`r aggr_mean_circ_age %>% filter(year == 2006, type == "MMC") %>% pull(upper)`
), and in 2020 it was
`r aggr_mean_circ_age %>% filter(year == 2020, type == "MMC") %>% pull(mean)`
(
`r aggr_mean_circ_age %>% filter(year == 2020, type == "MMC") %>% pull(lower)`
- 
`r aggr_mean_circ_age %>% filter(year == 2020, type == "MMC") %>% pull(upper)`
).
<!-- Therefore, again there wasn't a sizeable change in mean age at MMC. -->
Finally, the mean age at circumcision in 2006 for TMC was 
`r aggr_mean_circ_age %>% filter(year == 2006, type == "TMC") %>% pull(mean)`
(
`r aggr_mean_circ_age %>% filter(year == 2006, type == "TMC") %>% pull(lower)`
- 
`r aggr_mean_circ_age %>% filter(year == 2006, type == "TMC") %>% pull(upper)`
), and in 2020 it was
`r aggr_mean_circ_age %>% filter(year == 2020, type == "TMC") %>% pull(mean)`
(
`r aggr_mean_circ_age %>% filter(year == 2020, type == "TMC") %>% pull(lower)`
- 
`r aggr_mean_circ_age %>% filter(year == 2020, type == "TMC") %>% pull(upper)`.
So the mean age at circumcision for MC and MMC went up slightly from 2006 to 2020. 
However, it is important to note that the uncertainty ranges in our credible intervals 
overlap, so this increase could be seen as "non-significant", in some sense. 
In contrast, the mean age at circumcision for TMC was seen to decrease, altough again, 
not in a manor that exceeds our uncertainty bounds. 
This may suggest that VMMC implementation amongst 15-29 year olds has 
corresponded to an upward shift of mean age at circumcision for MC and MMC, and 
replaced some circumcisions which would previously have been performed as TMCs, 
which may explain the observed decrease in mean age of TMC.

As in the previous section, we can also compare how the mean age of circumcision differs 
amongst VMMC priority countries, versus their non-VMMC prority counterparts in SSA.

For VMMC priority countries, the mean age at circumcision for MC in 2006 was 
`r aggr_mean_circ_age_vmmc %>% filter(year == 2006, type == "MC") %>% pull(mean)`
(
`r aggr_mean_circ_age_vmmc %>% filter(year == 2006, type == "MC") %>% pull(lower)`
- 
`r aggr_mean_circ_age_vmmc %>% filter(year == 2006, type == "MC") %>% pull(upper)`
), while in 2020 it was
`r aggr_mean_circ_age_vmmc %>% filter(year == 2020, type == "MC") %>% pull(mean)`
(
`r aggr_mean_circ_age_vmmc %>% filter(year == 2020, type == "MC") %>% pull(lower)`
- 
`r aggr_mean_circ_age_vmmc %>% filter(year == 2020, type == "MC") %>% pull(upper)`
).
Similarly, for MMC in 2006 the mean age at circumcision was
`r aggr_mean_circ_age_vmmc %>% filter(year == 2006, type == "MMC") %>% pull(mean)`
(
`r aggr_mean_circ_age_vmmc %>% filter(year == 2006, type == "MMC") %>% pull(lower)`
- 
`r aggr_mean_circ_age_vmmc %>% filter(year == 2006, type == "MMC") %>% pull(upper)`
), while in 2020 it was
`r aggr_mean_circ_age_vmmc %>% filter(year == 2020, type == "MMC") %>% pull(mean)`
(
`r aggr_mean_circ_age_vmmc %>% filter(year == 2020, type == "MMC") %>% pull(lower)`
- 
`r aggr_mean_circ_age_vmmc %>% filter(year == 2020, type == "MMC") %>% pull(upper)`
).
<!-- Therefore, again there wasn't a sizeable change in mean age at MMC. -->
Lastly, the mean age at circumcision in 2006 for TMC was 
`r aggr_mean_circ_age_vmmc %>% filter(year == 2006, type == "TMC") %>% pull(mean)`
(
`r aggr_mean_circ_age_vmmc %>% filter(year == 2006, type == "TMC") %>% pull(lower)`
- 
`r aggr_mean_circ_age_vmmc %>% filter(year == 2006, type == "TMC") %>% pull(upper)`
), and in 2020 it was
`r aggr_mean_circ_age_vmmc %>% filter(year == 2020, type == "TMC") %>% pull(mean)`
(
`r aggr_mean_circ_age_vmmc %>% filter(year == 2020, type == "TMC") %>% pull(lower)`
- 
`r aggr_mean_circ_age_vmmc %>% filter(year == 2020, type == "TMC") %>% pull(upper)`.
So the mean age at circumcision for all circumcision types is higher in VMMC 
priority countries than overall in SSA, highlighting the historically low levels of MC and
TMC in these countries. 
Like in SSA overall, the mean age at circumcision for MC and MMC went up slightly, albeit not 
significantly when factoring in uncertainty bounds. 
Again, mean age at TMC also decreased in VMMC priority countries.

For non-VMMC priority countries, the mean age at MC in 2006 was 
`r aggr_mean_circ_age_non_vmmc %>% filter(year == 2006, type == "MC") %>% pull(mean)`
(
`r aggr_mean_circ_age_non_vmmc %>% filter(year == 2006, type == "MC") %>% pull(lower)`
- 
`r aggr_mean_circ_age_non_vmmc %>% filter(year == 2006, type == "MC") %>% pull(upper)`
), while in 2020 it was
`r aggr_mean_circ_age_non_vmmc %>% filter(year == 2020, type == "MC") %>% pull(mean)`
(
`r aggr_mean_circ_age_non_vmmc %>% filter(year == 2020, type == "MC") %>% pull(lower)`
- 
`r aggr_mean_circ_age_non_vmmc %>% filter(year == 2020, type == "MC") %>% pull(upper)`
).
Similarly, the mean age at MMC in 2006 was 
`r aggr_mean_circ_age_non_vmmc %>% filter(year == 2006, type == "MMC") %>% pull(mean)`
(
`r aggr_mean_circ_age_non_vmmc %>% filter(year == 2006, type == "MMC") %>% pull(lower)`
- 
`r aggr_mean_circ_age_non_vmmc %>% filter(year == 2006, type == "MMC") %>% pull(upper)`
), and in 2020 it was
`r aggr_mean_circ_age_non_vmmc %>% filter(year == 2020, type == "MMC") %>% pull(mean)`
(
`r aggr_mean_circ_age_non_vmmc %>% filter(year == 2020, type == "MMC") %>% pull(lower)`
- 
`r aggr_mean_circ_age_non_vmmc %>% filter(year == 2020, type == "MMC") %>% pull(upper)`
).
<!-- Therefore, again there wasn't a sizeable change in mean age at MMC. -->
Finally, the mean age at TMC in 2006 for TMC was 
`r aggr_mean_circ_age_non_vmmc %>% filter(year == 2006, type == "TMC") %>% pull(mean)`
(
`r aggr_mean_circ_age_non_vmmc %>% filter(year == 2006, type == "TMC") %>% pull(lower)`
- 
`r aggr_mean_circ_age_non_vmmc %>% filter(year == 2006, type == "TMC") %>% pull(upper)`
), and in 2020 it was
`r aggr_mean_circ_age_non_vmmc %>% filter(year == 2020, type == "TMC") %>% pull(mean)`
(
`r aggr_mean_circ_age_non_vmmc %>% filter(year == 2020, type == "TMC") %>% pull(lower)`
- 
`r aggr_mean_circ_age_non_vmmc %>% filter(year == 2020, type == "TMC") %>% pull(upper)`.
In contrast to our VMMC priority countries, these countries have a lower mean at 
circumcision for all circumcision types, largely driven by high TMC levels historically. 
It is interesting that the age at circumcision for MC and MMC actually increased for
non-VMMC priority countries, in contrast to SSA as a whole and the priority countries. 
Likewise, the mean age at TMC increased. 
This can perhaps be explained by a possible replacement of TMC with MMC in a 
traditional setting, so at generally younger ages than MMCs performed as part of 
VMMC programmes. 
This caused a natural shift upwards of the mean age at TMC, as younger individuals 
are now more likely to receive MMCs.

<!-- What is the narrative here?
- For SSA, MC and MMC age goes up, TMC goes down (done)
- For VMMC, average age for all substantially higher, but follows same pattern. 
  Could be argued that this is because people previously receiving TMCs are now receiving MMCs, 
  pushing mean age of TMC down further. 
- For non-VMMC, ages substantially lower. 
  MMC age actually decreases, suggesting MMC-nT occuring, replacing TMC! -->

## Decrease in TMC in SSA region Over Time

```{r}
x4 <- readRDS("data/inlines/04_inlines.RDS")
for (i in seq_along(x4)) {
  assign(names(x4)[i], x4[[i]])
}

```

Narrative here:
<!-- TMC coverage difference 2006 - 2020, SSA, VMMC, non-VMMC
- Decrease much more pronounced for non-VMMC
Number of TMCs performed in 2006 vs 2020, VMMC, non-VMMC
- Number increased more for VMMC countries (?)
Change for 15-29 vs 30-49
- Change relatively similar for non-VMMC, but 15-29 sees much greater change for 
  VMMC countries
Country with largest drop (CIV and ? in VMMC)
- Comment on population of CIV (28 mil) (?) -->

The average TMC coverage in SSA in 2006 was 
`r tmc_cov$mean_tmc_cov %>% filter(year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r tmc_cov$mean_tmc_cov %>% filter(year == 2006) %>% pull(lower) %>% to_percentage()`
to 
x
`r tmc_cov$mean_tmc_cov %>% filter(year == 2006) %>% pull(upper) %>% to_percentage()`
). 
This fell to 
`r tmc_cov$mean_tmc_cov %>% filter(year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r tmc_cov$mean_tmc_cov %>% filter(year == 2020) %>% pull(lower) %>% to_percentage()`
to 
x
`r tmc_cov$mean_tmc_cov %>% filter(year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020, a decrease of 
`r to_percentage(tmc_cov$mean_tmc_cov_diff$mean)`. 
In non-VMMC priority countries, TMC coverage was 
`r tmc_cov_non_vmmc$mean_tmc_cov %>% filter(year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r tmc_cov_non_vmmc$mean_tmc_cov %>% filter(year == 2006) %>% pull(lower) %>% to_percentage()`
to 
x
`r tmc_cov_non_vmmc$mean_tmc_cov %>% filter(year == 2006) %>% pull(upper) %>% to_percentage()`
). 
This fell to 
`r tmc_cov_non_vmmc$mean_tmc_cov %>% filter(year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r tmc_cov_non_vmmc$mean_tmc_cov %>% filter(year == 2020) %>% pull(lower) %>% to_percentage()`
to 
x
`r tmc_cov_non_vmmc$mean_tmc_cov %>% filter(year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020, representing a decrease of 
`r to_percentage(tmc_cov_non_vmmc$mean_tmc_cov_diff$mean)`
.
In contrast, TMC coverage in VMMC priority countries was 
`r tmc_cov_vmmc$mean_tmc_cov %>% filter(year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r tmc_cov_vmmc$mean_tmc_cov %>% filter(year == 2006) %>% pull(lower) %>% to_percentage()`
to 
x
`r tmc_cov_vmmc$mean_tmc_cov %>% filter(year == 2006) %>% pull(upper) %>% to_percentage()`
). 
This fell to 
`r tmc_cov_vmmc$mean_tmc_cov %>% filter(year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r tmc_cov_vmmc$mean_tmc_cov %>% filter(year == 2020) %>% pull(lower) %>% to_percentage()`
to 
x
`r tmc_cov_vmmc$mean_tmc_cov %>% filter(year == 2020) %>% pull(upper) %>% to_percentage()`
) in 2020, representing a decrease of 
`r to_percentage(tmc_cov_vmmc$mean_tmc_cov_diff$mean)`.
Therefore, the decrease in TMC found throughout SSA is much more pronounced for 
non-VMMC countries.

In 2006, 
`r tmcs_performed$tmcs_performed %>% filter(year == 2006) %>% pull(mean) %>% round(0)`
(
`r tmcs_performed$tmcs_performed %>% filter(year == 2006) %>% pull(lower) %>% round(0)`
- 
`r tmcs_performed$tmcs_performed %>% filter(year == 2006) %>% pull(upper) %>% round(0)`
) TMCs were performed, while in 2020 this increased to 
`r tmcs_performed$tmcs_performed %>% filter(year == 2020) %>% pull(mean) %>% round(0)`
(
`r tmcs_performed$tmcs_performed %>% filter(year == 2020) %>% pull(lower) %>% round(0)`
-
`r tmcs_performed$tmcs_performed %>% filter(year == 2020) %>% pull(upper) %>% round(0)`
), representing a 
`r round(-tmcs_performed$tmcs_performed_diff$mean, 0)`
increase in annual TMCs.
However, this increase must be taken within the context of a large increase
in population in SSA over this period, from 1.23 billion in 2006 to 1.80 billion in 2020, 
and as seen in our coverage estimates above, constitutes a decrease in the 
number of TMCs performed per capita in the region (need to calculate this?!)
<!-- populations %>% filter(age_group == "0+", year %in% spec_years) %>% group_by(year) %>% summarise(population = sum(population, na.rm = TRUE)) -->
<!-- populations %>% filter(age_group == "0+", year %in% spec_years, iso3 %in% vmmc_iso3) %>% group_by(year) %>% summarise(population = sum(population, na.rm = TRUE)) -->
<!-- populations %>% filter(age_group == "0+", year %in% spec_years, !(iso3 %in% vmmc_iso3)) %>% group_by(year) %>% summarise(population = sum(population, na.rm = TRUE)) -->
In non-VMMC priority countries, the annual number of TMCs performed changed from 
`r tmcs_performed_non_vmmc$tmcs_performed %>% filter(year == 2006) %>% pull(mean) %>% round(0)`
(
`r tmcs_performed_non_vmmc$tmcs_performed %>% filter(year == 2006) %>% pull(lower) %>% round(0)`
- 
`r tmcs_performed_non_vmmc$tmcs_performed %>% filter(year == 2006) %>% pull(upper) %>% round(0)`
) in 2006 to 
`r tmcs_performed_non_vmmc$tmcs_performed %>% filter(year == 2020) %>% pull(mean) %>% round(0)`
(
`r tmcs_performed_non_vmmc$tmcs_performed %>% filter(year == 2020) %>% pull(lower) %>% round(0)`
- 
`r tmcs_performed_non_vmmc$tmcs_performed %>% filter(year == 2020) %>% pull(upper) %>% round(0)`
) in 2020, or by 
`r round(-tmcs_performed_non_vmmc$tmcs_performed_diff$mean, 0)`
.
This is within the context of the population of non-VMMC priority countries increasing 
from 749 million to 1.11 billion from 2006 to 2020. 
In VMMC priority countries, the number of TMCs performed went from 
`r tmcs_performed_vmmc$tmcs_performed %>% filter(year == 2006) %>% pull(mean) %>% round(0)`
(
`r tmcs_performed_vmmc$tmcs_performed %>% filter(year == 2006) %>% pull(lower) %>% round(0)`
- 
`r tmcs_performed_vmmc$tmcs_performed %>% filter(year == 2006) %>% pull(upper) %>% round(0)`
) in 2006 to 
`r tmcs_performed_vmmc$tmcs_performed %>% filter(year == 2006) %>% pull(mean) %>% round(0)`
(
`r tmcs_performed_vmmc$tmcs_performed %>% filter(year == 2006) %>% pull(lower) %>% round(0)`
- 
`r tmcs_performed_vmmc$tmcs_performed %>% filter(year == 2006) %>% pull(upper) %>% round(0)`
) in 2020, a fall of 
`r round(-tmcs_performed_vmmc$tmcs_performed_diff$mean, 0)`.
In VMMC priority countries, the population increased from 481 million to 694 million. 
TODO: Comment on this

<!-- TODO: Change to 15-29, right? -->
We can also compare the differences in TMC coverage change between 15-29 and 
30-49 year olds, in SSA as a whole and in VMMC and non-VMMC priority countries.

For the 15-29 age group, TMC coverage was reduced from 
`r tmc_cov_ssa_15_29_30_49[[1]] %>% filter(year == 2006, age_group == "15-29") %>% pull(mean) %>% to_percentage()`
(
`r tmc_cov_ssa_15_29_30_49[[1]] %>% filter(year == 2006, age_group == "15-29") %>% pull(lower) %>% to_percentage()`
to 
`r tmc_cov_ssa_15_29_30_49[[1]] %>% filter(year == 2006, age_group == "15-29") %>% pull(upper) %>% to_percentage()`
) in 2006 to 
`r tmc_cov_ssa_15_29_30_49[[1]] %>% filter(year == 2020, age_group == "15-29") %>% pull(mean) %>% to_percentage()`
(
`r tmc_cov_ssa_15_29_30_49[[1]] %>% filter(year == 2020, age_group == "15-29") %>% pull(lower) %>% to_percentage()`
to 
`r tmc_cov_ssa_15_29_30_49[[1]] %>% filter(year == 2020, age_group == "15-29") %>% pull(upper) %>% to_percentage()`
) in 2020, a drop of 
`r tmc_cov_ssa_15_29_30_49[[2]] %>% filter(age_group == "15-29") %>% pull(mean) %>% to_percentage()`
.
This contrasted with the 30-49 age group, where TMC coverage went from 
`r tmc_cov_ssa_15_29_30_49[[1]] %>% filter(year == 2006, age_group == "30-49") %>% pull(mean) %>% to_percentage()`
(
`r tmc_cov_ssa_15_29_30_49[[1]] %>% filter(year == 2006, age_group == "30-49") %>% pull(lower) %>% to_percentage()`
to 
`r tmc_cov_ssa_15_29_30_49[[1]] %>% filter(year == 2006, age_group == "30-49") %>% pull(upper) %>% to_percentage()`
) in 2006 to 
`r tmc_cov_ssa_15_29_30_49[[1]] %>% filter(year == 2020, age_group == "30-49") %>% pull(mean) %>% to_percentage()`
(
`r tmc_cov_ssa_15_29_30_49[[1]] %>% filter(year == 2020, age_group == "30-49") %>% pull(lower) %>% to_percentage()`
to 
`r tmc_cov_ssa_15_29_30_49[[1]] %>% filter(year == 2020, age_group == "30-49") %>% pull(upper) %>% to_percentage()`
) in 2020, a drop of 
`r tmc_cov_ssa_15_29_30_49[[2]] %>% filter(age_group == "30-49") %>% pull(mean) %>% to_percentage()`
.
Therefore, in SSA as a whole TMC coverage decreased slightly more amongst 15-29 
year olds than in 30-49 year olds, likely as a result of VMMC programme focus on 
this age group.

For the 15-29 year olds in non-VMMC priority countries, TMC coverage was reduced from 
`r tmc_cov_non_vmmc_15_29_30_49[[1]] %>% filter(year == 2006, age_group == "15-29") %>% pull(mean) %>% to_percentage()`
(
`r tmc_cov_non_vmmc_15_29_30_49[[1]] %>% filter(year == 2006, age_group == "15-29") %>% pull(lower) %>% to_percentage()`
to 
`r tmc_cov_non_vmmc_15_29_30_49[[1]] %>% filter(year == 2006, age_group == "15-29") %>% pull(upper) %>% to_percentage()`
) in 2006 to 
`r tmc_cov_non_vmmc_15_29_30_49[[1]] %>% filter(year == 2020, age_group == "15-29") %>% pull(mean) %>% to_percentage()`
(
`r tmc_cov_non_vmmc_15_29_30_49[[1]] %>% filter(year == 2020, age_group == "15-29") %>% pull(lower) %>% to_percentage()`
to 
`r tmc_cov_non_vmmc_15_29_30_49[[1]] %>% filter(year == 2020, age_group == "15-29") %>% pull(upper) %>% to_percentage()`
) in 2020, a drop of 
`r tmc_cov_non_vmmc_15_29_30_49[[2]] %>% filter(age_group == "15-29") %>% pull(mean) %>% to_percentage()`
.
This contrasted with the 30-49 age group, where TMC coverage went from 
`r tmc_cov_non_vmmc_15_29_30_49[[1]] %>% filter(year == 2006, age_group == "30-49") %>% pull(mean) %>% to_percentage()`
(
`r tmc_cov_non_vmmc_15_29_30_49[[1]] %>% filter(year == 2006, age_group == "30-49") %>% pull(lower) %>% to_percentage()`
to 
`r tmc_cov_non_vmmc_15_29_30_49[[1]] %>% filter(year == 2006, age_group == "30-49") %>% pull(upper) %>% to_percentage()`
) in 2006 to 
`r tmc_cov_non_vmmc_15_29_30_49[[1]] %>% filter(year == 2020, age_group == "30-49") %>% pull(mean) %>% to_percentage()`
(
`r tmc_cov_non_vmmc_15_29_30_49[[1]] %>% filter(year == 2020, age_group == "30-49") %>% pull(lower) %>% to_percentage()`
to 
`r tmc_cov_non_vmmc_15_29_30_49[[1]] %>% filter(year == 2020, age_group == "30-49") %>% pull(upper) %>% to_percentage()`
) in 2020, a drop of 
`r tmc_cov_non_vmmc_15_29_30_49[[2]] %>% filter(age_group == "30-49") %>% pull(mean) %>% to_percentage()`
.
Therefore in non-VMMC priority countries the decrease in TMC coverage was larger
for both age groups than in SSA as a whole, but the difference between this decrease
for 15-29 year olds versus 30-49 year olds is less pronounced.

For VMMC priority countries, the 15-29 age group, TMC coverage was reduced from 
`r tmc_cov_vmmc_15_29_30_49[[1]] %>% filter(year == 2006, age_group == "15-29") %>% pull(mean) %>% to_percentage()`
(
`r tmc_cov_vmmc_15_29_30_49[[1]] %>% filter(year == 2006, age_group == "15-29") %>% pull(lower) %>% to_percentage()`
to 
`r tmc_cov_vmmc_15_29_30_49[[1]] %>% filter(year == 2006, age_group == "15-29") %>% pull(upper) %>% to_percentage()`
) in 2006 to 
`r tmc_cov_vmmc_15_29_30_49[[1]] %>% filter(year == 2020, age_group == "15-29") %>% pull(mean) %>% to_percentage()`
(
`r tmc_cov_vmmc_15_29_30_49[[1]] %>% filter(year == 2020, age_group == "15-29") %>% pull(lower) %>% to_percentage()`
to 
`r tmc_cov_vmmc_15_29_30_49[[1]] %>% filter(year == 2020, age_group == "15-29") %>% pull(upper) %>% to_percentage()`
) in 2020, representing a fall of 
`r tmc_cov_vmmc_15_29_30_49[[2]] %>% filter(age_group == "15-29") %>% pull(mean) %>% to_percentage()`
.
In the 30-49 age group, TMC coverage in VMMC priority countries dropped from 
`r tmc_cov_vmmc_15_29_30_49[[1]] %>% filter(year == 2006, age_group == "30-49") %>% pull(mean) %>% to_percentage()`
(
`r tmc_cov_vmmc_15_29_30_49[[1]] %>% filter(year == 2006, age_group == "30-49") %>% pull(lower) %>% to_percentage()`
to 
`r tmc_cov_vmmc_15_29_30_49[[1]] %>% filter(year == 2006, age_group == "30-49") %>% pull(upper) %>% to_percentage()`
) in 2006 to 
`r tmc_cov_vmmc_15_29_30_49[[1]] %>% filter(year == 2020, age_group == "30-49") %>% pull(mean) %>% to_percentage()`
(
`r tmc_cov_vmmc_15_29_30_49[[1]] %>% filter(year == 2020, age_group == "30-49") %>% pull(lower) %>% to_percentage()`
to 
`r tmc_cov_vmmc_15_29_30_49[[1]] %>% filter(year == 2020, age_group == "30-49") %>% pull(upper) %>% to_percentage()`
) in 2020, a drop of 
`r tmc_cov_vmmc_15_29_30_49[[2]] %>% filter(age_group == "30-49") %>% pull(mean) %>% to_percentage()`
.
Therefore, in VMMC priority countries there is a much greater disparity in the 
drop in TMC coverage between 15-29 and 30-49 year olds. 15-29 year olds saw 
a much greater drop in TMC coverage, likely resulting from VMMC programme focus on 
these ages.

The largest decrease in TMC coverage from 2006 to 2020 was in 
`r to_cntry_name(spec_country_tmc_decrease$mean_tmc_cov_diff$iso3)`, where TMC coverage
fell from 
`r spec_country_tmc_decrease[[1]] %>% filter(year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r spec_country_tmc_decrease[[1]] %>% filter(year == 2006) %>% pull(lower) %>% to_percentage()`
 to 
`r spec_country_tmc_decrease[[1]] %>% filter(year == 2006) %>% pull(upper) %>% to_percentage()`
) to 
`r spec_country_tmc_decrease[[1]] %>% filter(year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r spec_country_tmc_decrease[[1]] %>% filter(year == 2020) %>% pull(lower) %>% to_percentage()`
 to 
`r spec_country_tmc_decrease[[1]] %>% filter(year == 2020) %>% pull(upper) %>% to_percentage()`, a 
decrease of 
`r to_percentage(spec_country_tmc_decrease[[2]]$mean)`
.

The largest decrease in TMC coverage amongst VMMC countries was from 
`r to_cntry_name(spec_country_tmc_decrease$mean_tmc_cov_diff$iso3)`, where TMC coverage
fell from 
`r spec_country_tmc_decrease[[1]] %>% filter(year == 2006) %>% pull(mean) %>% to_percentage()`
(
`r spec_country_tmc_decrease[[1]] %>% filter(year == 2006) %>% pull(lower) %>% to_percentage()`
 to 
`r spec_country_tmc_decrease[[1]] %>% filter(year == 2006) %>% pull(upper) %>% to_percentage()`
) to 
`r spec_country_tmc_decrease[[1]] %>% filter(year == 2020) %>% pull(mean) %>% to_percentage()`
(
`r spec_country_tmc_decrease[[1]] %>% filter(year == 2020) %>% pull(lower) %>% to_percentage()`
 to 
`r spec_country_tmc_decrease[[1]] %>% filter(year == 2020) %>% pull(upper) %>% to_percentage()`, a 
decrease of 
`r to_percentage(spec_country_tmc_decrease[[2]]$mean)`
.

TODO: Could possibly add smallest decrease here? Would be more interesting for 
non-VMMC countries, could show where MMC has not begun to replace TMC
Could also have number of countries in both regions which have experienced a 
drop in TMC coverage?? Might be interesting for non-VMMC

## UNAIDS Goals Progress amongst 10-29 Year Olds in VMMC Target Countries

```{r}
# x5 <- readRDS("data/inlines/05_inlines.RDS")
# for (i in seq_along(x5)) {
#   assign(names(x5)[i], x5[[i]])
# }
# aggr_mean_circ_age <- tidy_age(aggr_mean_circ_age)
# aggr_mean_circ_age_non_vmmc <- tidy_age(aggr_mean_circ_age_non_vmmc)
# aggr_mean_circ_age_vmmc <- tidy_age(aggr_mean_circ_age_vmmc)
```

Amongst the 13 VMMC priority countries modelled, just 
x
 countries were projected to have achieved the VMMC target of 90\% MC coverage amongst 10-29 year olds, 
x
 and 
x
. 
MC coverage across these countries increased from 
x
\% (
x
\% to 
x
\%) in 2006 to 
x
\% (
x
\% to 
x
\%) in 2020.

x
 (
x
 to 
x
) circumcisions were performed between 2007 and 2020 in these countries. 
Of these, 
x
 (
x
 to 
x
) were MMCs. 
No countries were e
x
pected to have reached this goal in all districts. 
The number of annual MCs increased from 
x
 (
x
 to 
x
) in 2006 to 
x
 (
x
 to 
x
) in 2020.

x
 (
x
 to 
x
) additional circumcisions are required to achieve 90\% MC coverage in all VMMC countries, 
x
 (
x
 to 
x
) times the projected number of MCs for 2020. 
The largest increase in MC coverage amongst VMMC priority countries was in 
x
, from 
x
\% (
x
\% to 
x
\%) in 2006 to 
x
\% (
x
\% to 
x
\%) in 2020, a 
x
\% (
x
\% to 
x
\%) increase. 
In contrast, the lowest increase in MC coverage in these countries was in 
x
, from 
x
\% (
x
\% to 
x
\%) in 2006 to 
x
\% (
x
\% to 
x
\%) in 2020, a 
x
\% (
x
\% to 
x
\%) increase.

However, these figures belie significant subnational heterogeneity in MC coverage amongst VMMC prioirty countries.

x
/
x
 districts in these countries were e
x
pected to have reached 90\% MC coverage amongst 10-29 year olds in 2020, up from 
x
/
x
 in 2006, before VMMC programmes began. 
Even in 
x
, where MC was projected to be lowest in 2020, 
x
/
x
 districts were forecasted to have reached 90\% MC coverage for men aged 10-29 years old. 
MC coverage there increased from 
x
(
x
to 
x
) in 2006 to 
x
(
x
to 
x
) in 2020, a 
x
(
x
to 
x) increase.